---
title: "Eval Tree Analysis Code"
author: "Bobea"
date: "2024-02-03"
output: html_document
---
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      TreeAlert Solar Wind Example                         
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

load data
```{r, include=FALSE, warning=FALSE, message=FALSE}
# 2. Load the four specific files shown in the screenshot
# # We use file.path to ensure the slashes are handled correctly for Windows
# load(file.path(base_path, "train_solar_v2.RData"), verbose = TRUE)
# load(file.path(base_path, "train_wind_v2.RData"), verbose = TRUE)
# load(file.path(base_path, "test_wind_v2.RData"), verbose = TRUE)
# load(file.path(base_path, "test_solar_v2.RData"), verbose = TRUE)
```
Libraries
```{r, include=FALSE, warning=FALSE, message=FALSE}
options(scipen = 999)
library(readr); library(tidyr); library(feasts); library(scoringRules) # For Data
library(tsibble); library(fable); library(fabletools) # For Forecasting
library(dplyr); library(lubridate) # For Data Wrangling
library(ggplot2); library(ggtext); library(grid); library(gridExtra); # Plot
library(slider) # Plot
library(scales); library(rpart);  library(rpart.plot);
```
Define Functions:
```{r}
temporal_encoding <- function(data) {
  # Convert Time column to Date-Time format and create temporal features
   data <- data %>%
    mutate(
      Time = parse_date_time(Time, orders = c("ymd HMS", "ymd HM", "ymd H", "ymd")),
      Hour = factor(hour(Time)),
      Minute = factor(minute(Time)),
      Weekday = factor(lubridate::wday(Time, label = TRUE)),
      DayofMonth = factor(mday(Time)),
      WeekofMonth = factor(ceiling(day(Time) / 7)),
      Month = factor(lubridate::month(Time, label = TRUE))
    ) %>%
    arrange(Time)

 # Number of hours in 1–4 weeks
week_lags <- c(168, 336, 504, 672)

for (i in seq_along(week_lags)) {
  lag_name <- paste0("Lag_", i, "_Week")
  data[[lag_name]] <- dplyr::lag(data$.resid, n = week_lags[i])
}

  # Fill NAs in all lag columns with previous value
  data <- data %>%
    arrange(Time) %>%
    fill(starts_with("Lag_"), .direction = "down")

  return(data)
}

# process_data converts a random df into a format usable by Tree Alert!
process_data <- function(data, time_col, actual_col, forecast_col) {

  data %>%
    # Select specified columns and rename them
    #dplyr::select({{ time_col }}, {{ actual_col }}, {{ forecast_col }}) %>%
    dplyr::rename(
      Time = {{ time_col }},         # Rename timestamp
      actual = {{ actual_col }},     # Rename actual load
      forecast = {{ forecast_col }}  # Rename forecast load
    ) %>%
    # Convert to datetime and handle multiple possible formats
    dplyr::mutate(
      Time = parse_date_time(Time, orders = c("ymd HMS", "ymd HM", "mdy HMS", "mdy HM", "dmy HMS", "dmy HM")),
      .resid = winkler_score     #  for forecast error
    ) %>%
    na.omit()                        # Remove any NA values
}

# To generate geom_point layers for 'All Filters Actual Vs Predicted'
add_fitted_points_func <- function(data_list, color_list, shape_values) {
  layers <- lapply(seq_along(data_list), function(i) {
    geom_point(data = data_list[[i]], 
               aes(x = Time, y = Value, shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
               size = 2.5, stroke = 1, color = color_list[i])
  })
  return(layers)
}

# To generate geom_point layers for 'All Filters Residuals'
add_resid_points_func <- function(data_list, color_list, shape_values) {
  layers <- lapply(seq_along(data_list), function(i) {
    geom_point(data = data_list[[i]], 
               aes(x = Time, y = .resid, shape = .resid >= top_5_percent_value | 
                     .resid <= bottom_5_percent_value),
               size = 2.5, stroke = 1, color = color_list[i])
  })
  return(layers)
}

# Function to create the plot for a specific month
create_monthly_plot <- function(data, start_date, end_date) {
  plot <- ggplot(data, aes(x = Time)) +
    geom_line(aes(y = Value, color = "Actual"), show.legend = FALSE) +
    geom_line(aes(y = model_1, color = "Fitted"), alpha = 0.6, show.legend = FALSE) +
    # changed to skip rules 1-2 
    add_fitted_points_func(list_of_dfs[start_row:length(list_of_dfs)], 
                       color_string[start_row:length(list_of_dfs)], 
                       c("TRUE" = 1, "FALSE" = 16))
    #add_fitted_points_func(list_of_dfs, color_string, # make geom_point layers
    #                                  c("TRUE" = 1, "FALSE" = 16)) + 
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
    scale_color_manual(values = c("Actual" = "#4682B4", "Fitted" = "#FF8C00")) +
    scale_x_datetime(limits = as.POSIXct(c(start_date, end_date)), 
                     date_labels = "%m-%d",
                     #date_breaks = "3 days",
                     expand = c(0.02, 0.02)) +
    labs(title = paste(""), x = NULL, y = NULL, color = "Variable") +
    theme_bw() +
    theme(legend.position = "none",
          plot.title = element_markdown(size = 10),
          plot.title.position = "panel")
  
  return(plot)
}

# -------------------- Lift Functions --------------------
compute_lift_numeric <- function(df, actual_col, pred_col, n_groups = 20) {
  # Lift is calculated using all data, but only returns lift for the first group  
  actual <- abs(df[[actual_col]])
  score  <- abs(df[[pred_col]])
  
  # Sort by predicted risk
  ord        <- order(score, decreasing = TRUE)
  actual_ord <- actual[ord]
  n          <- length(actual_ord)
  
  # Ventile grouping
  group <- ceiling(seq_len(n) / (n / n_groups))
  group[group > n_groups] <- n_groups
  
  df_lift <- data.frame(group = group, error = actual_ord) %>%
    dplyr::group_by(group) %>%
    dplyr::summarise(meanError = mean(error, na.rm = TRUE), .groups = "drop")
  
  # Normalize
  df_lift$meanError <- df_lift$meanError / mean(actual_ord, na.rm = TRUE)
  
  # Return first ventile lift
  df_lift$meanError[df_lift$group == 1]
}

make_lift_charts <- function(df, actual_col, pred_col, title_prefix = "Model") {
  
  actual <- abs(df[[actual_col]])
  score  <- abs(df[[pred_col]])
  
  # Sort by predicted risk
  ord <- order(score, decreasing = TRUE)
  actual_ord <- actual[ord]
  n <- length(actual_ord)
  
  # Cumulative gains
  df_cum <- data.frame(
    ncases  = c(0, seq_len(n)),
    cumError = c(0, cumsum(actual_ord))
  )
  
  # Ventiles
  n_groups <- 20
  group <- ceiling(seq_len(n) / (n / n_groups))
  group[group > n_groups] <- n_groups
  
  df_lift <- data.frame(
    group = group,
    error = actual_ord
  ) %>%
    dplyr::group_by(group) %>%
    dplyr::summarise(meanError = mean(error), .groups = "drop")
  
  df_lift$percentile <- df_lift$group * 5
  df_lift$meanError  <- df_lift$meanError / mean(actual_ord)
  
  g1 <- ggplot(df_cum, aes(ncases, cumError)) +
    geom_line(color = "#4682B4") +
    geom_abline(slope = sum(actual_ord)/n, intercept = 0,
                linetype = "dashed", color = "gray") +
    labs(title = paste(title_prefix, "– Cumulative Gains"),
         x = "# Cases", y = "Cumulative Error") +
    theme_minimal()
  
  g2 <- ggplot(df_lift, aes(percentile, meanError)) +
    geom_col(fill = "#FF8C00") +
    labs(title = paste(title_prefix, "– Ventile Lift"),
         x = "Percentile", y = "Ventile Mean / Global Mean") +
    theme_minimal()
  
  gridExtra::grid.arrange(g1, g2, ncol = 2)
  
  df_lift$meanError[df_lift$percentile == 5]
}

# --Compute lift-- 
compute_lift <- function(data, model, pred_col_name, label, plot = TRUE, n_groups = 20) {

  # Remove existing prediction column if present
  data <- data %>% dplyr::select(-dplyr::any_of(pred_col_name))

  # Predict leaf-node values (force numeric)
  leaf_nodes <- as.numeric(predict(model, newdata = data, type = "vector"))

  # Attach predictions
  data_with_pred <- dplyr::bind_cols(
    data,
    !!pred_col_name := leaf_nodes
  )

  # Plot or numeric lift
  lift_value <- if (plot) {
    make_lift_charts(
      df           = data_with_pred,
      actual_col   = ".resid",
      pred_col     = pred_col_name,
      title_prefix = label
    )
  } else {
    compute_lift_numeric(
      df         = data_with_pred,
      actual_col = ".resid",
      pred_col   = pred_col_name,
      n_groups   = n_groups
    )
  }

  cat(label, "Lift (First Ventile):", lift_value, "\n")

  return(list(
    lift        = lift_value,
    merged_data = data_with_pred
  ))
}

# -------------------- Prep Parameters --------------------
color_string <- c("#9400D3", "#008000", "#0000FF", "#FF8C00", "#000066",
                  "#FF00FF", "#AB0000", "#FF69B4", "#7B68EE", "#Ff0000", 
                  "#7B68EE", "#20B2AA", "#703366", "#32CD32", "#CC6633", 
                  "#C77CFF", "#8A2BE2", "#00FF7F", "#00BFFF", "#FFA500", 
                  "#66CDAA", "#BA55D3", "#00CED1", "#4682B4", "#00FFFF")
```
Load files 
```{r, warning=FALSE}
 base_path <- "C:/Users/Kagebi/Documents/Research/TreeAlert and Lift/solar_wind/"
  load(file.path(base_path, "all_wind_merged.RData"), verbose = TRUE)

# Select only the timestamp columns (1-2) and your specified columns using dplyr
raw_data <- all_wind_data
```
# START!  
Select your data!
```{r}
#Format the timestamp to ISO 8601 (T and Z format)
# This converts the POSIXct object into the character string format your raw_data uses
all_wind_data[, valid_datetime := format(valid_datetime, format = "%Y-%m-%dT%H:%M:%SZ")]

# # Define the specific column names you want to keep
# Actual_column <- "actual"
# Forecast_column <- "forecast"
# time_col = "valid_datetime"
# 
# 
# data_start <-"2020-09-20 00:30:00" 
# data_end <-"2023-06-01 23:30:00"
# 
# train_start <- "2020-09-20 00:30:00"
# train_end <- "2023-02-01 00:00:00"
# 
# test_start <- train_end 
# test_end<- data_end 
# 
# #How many rules? 
# TOP_K_RULES <- 5

# 6-Month Window (Keep Start)
# ----------------------------
data_start <- "2020-09-01 00:00:00"
data_end   <- "2021-03-01 00:00:00"

# ~80/20 split (5 months / 1 month)
train_start <- data_start
train_end   <- "2021-02-01 00:00:00"

test_start  <- train_end
test_end    <- data_end

# Number of top rules
TOP_K_RULES <- 5
#--------------------------------------------------------

data_df <- process_data(
  raw_data, 
  time_col = time_col, 
  actual_col = Actual_column, 
  forecast_col = Forecast_column
)

# Reduce data size with a Time Filter (Optional)
data_df <- data_df[data_df$Time >= as.POSIXct(data_start, tz = "UTC"), ]
data_df <- data_df[data_df$Time < as.POSIXct(data_end, tz = "UTC"), ]
```
Create a new data frame that matches the structure of data_df
```{r}
# Aggregate by hour for more clear plots
data_filtered <- data_df %>%
  mutate(hour = format(Time, "%Y-%m-%d %H")) %>%
  group_by(hour) %>%
  filter(actual == max(actual)) %>%
  ungroup() %>%
  select(-hour) 

data_df <- temporal_encoding(data_filtered) 

# Create a copy with the required columns and modifications
data_df <- data_df %>%
  mutate(
    Value = actual,
    .fitted = forecast,
    .resid = .resid
  ) %>%
  select(-actual, -forecast) %>%
  as_tibble()  # Convert to tibble if consistency is needed

#------------------------------------
#Data Post Processing + ~~ TRAIN TEST SPLIT ~~

#option 2 --  use time to split
train_set <- data_df[data_df$Time > as.POSIXct(train_start, tz = "UTC") & 
                    data_df$Time < as.POSIXct(train_end, tz = "UTC"),]
test_set <- data_df[data_df$Time >= as.POSIXct(test_start, tz = "UTC") & 
                    data_df$Time < as.POSIXct(test_end, tz = "UTC"),]


# cap Outliers (top only)
cap_99 <- quantile(train_set$.resid, 0.99, na.rm = TRUE)

train_set$.resid <- pmin(train_set$.resid, cap_99)
test_set$.resid  <- pmin(test_set$.resid,  cap_99)
data_df$.resid   <- pmin(data_df$.resid,   cap_99)

#------------Top bottom 5% of data tree ------------
# Calculate quantiles
top_5_percent <- quantile(train_set$.resid, probs = 0.95, na.rm = TRUE)
bottom_5_percent <- quantile(train_set$.resid, probs = 0.05, na.rm = TRUE)
# Extract values
top_5_percent_value <- top_5_percent[[1]] 
bottom_5_percent_value <- bottom_5_percent[[1]]
```
# Make Trees & Pull Rules
```{r}
#------------Regression tree on .resid + temporal variables------------
tree_model <- rpart(.resid ~ Minute + Hour + Weekday,
                    data = train_set, method = "anova", 
                    cp = 0.0000, minbucket = 10, maxdepth = 10)
# cp = 0.000, minbucket = 3
# cp = 0.00006, minbucket = 2
# + DayofMonth+ WeekofMonth, + Lag_2_Week + Minute +
tree_rules <- rpart.rules(tree_model, cover = FALSE, digits = 4, nn = FALSE)

#Attached the predicted tree nodes to DF and Maps rules/leafs to Observations
leaf_nodes <- predict(tree_model, newdata = data_df, type = "vector")
heart_merged_with_leaf <- dplyr::bind_cols(data_df, Leaf_Node = leaf_nodes)

#Lift Analysis
# update lift groupings.  e.g. decile, ventile, ect. 
N_GROUPS <- 20

# Calc lift
train_results <- compute_lift(train_set, tree_model, "Leaf_Node","Baseline (Train)", 
                              plot = TRUE, n_groups = N_GROUPS)
test_results  <- compute_lift(test_set,  tree_model, "Leaf_Node", "Baseline (Test)",
                              plot = TRUE, n_groups = N_GROUPS)
full_results  <- compute_lift(data_df,   tree_model, "Leaf_Node", "Baseline (Full)",  
                              plot = TRUE, n_groups = N_GROUPS)
```

## Grid Search: Joint Optimization of Tree Parameters + Predictor Variables  ##
```{r}
## Grid Search: Sensitivity Analysis + Full Logging  ##

library(tibble)

# ------------------------------------------------------------------
# 1. Initialize results table
# ------------------------------------------------------------------
sensitivity_results <- tibble(
  run_id           = integer(),
  cp               = numeric(),
  minbucket        = integer(),
  maxdepth         = integer(),
  vars             = character(),
  train_lift       = numeric(),
  test_lift        = numeric(),
  full_lift        = numeric(),
  train_precision  = numeric(),
  test_precision   = numeric(),
  train_flagged    = integer(),
  test_flagged     = integer(),
  train_tp         = integer(),
  train_fp         = integer(),
  test_tp          = integer(),
  test_fp          = integer()
)

# ------------------------------------------------------------------
# 2. Candidate predictors + grid
# ------------------------------------------------------------------


param_grid <- expand.grid(
  cp        = c(0, 0.0001, 0.001),
  minbucket = c(5, 10, 20),
  maxdepth  = c(5, 10, 15),
  vars      = var_combinations,
  stringsAsFactors = FALSE
)
param_grid$vars <- "Hour + Weekday + Month"

# ------------------------------------------------------------------
# 3. Grid search loop with TOP-K RULE EVALUATION
# ------------------------------------------------------------------

for (i in seq_len(nrow(param_grid))) {

  params <- param_grid[i, ]
  fmla   <- .resid ~ Hour + Weekday + Month

  model <- rpart(
    fmla,
    data   = train_set,
    method = "anova",
    control = rpart.control(
      cp        = params$cp,
      minbucket = params$minbucket,
      maxdepth  = params$maxdepth
    )
  )

  # ------------------------------------------------------------
  # LIFT (unchanged – whole-tree ranking quality)
  # ------------------------------------------------------------
  train_lift <- compute_lift_numeric(
    df = bind_cols(train_set, Leaf_Node = predict(model, train_set)),
    actual_col = ".resid",
    pred_col   = "Leaf_Node"
  )

  test_lift <- compute_lift_numeric(
    df = bind_cols(test_set, Leaf_Node = predict(model, test_set)),
    actual_col = ".resid",
    pred_col   = "Leaf_Node"
  )

  full_lift <- compute_lift_numeric(
    df = bind_cols(data_df, Leaf_Node = predict(model, data_df)),
    actual_col = ".resid",
    pred_col   = "Leaf_Node"
  )

  # ------------------------------------------------------------
  # STEP 2: Extract TOP-K highest-risk LEAF RULES (SCALABLE)
  # ------------------------------------------------------------
  train_tmp <- bind_cols(
    train_set,
    Leaf_Node = predict(model, train_set)
  )

  leaf_summary <- train_tmp %>%
    group_by(Leaf_Node) %>%
    summarise(
      mean_resid = mean(.resid, na.rm = TRUE),
      .groups = "drop"
    )

  top_k_rules <- leaf_summary %>%
    arrange(desc(abs(mean_resid))) %>%
    slice(1:TOP_K_RULES) %>%
    pull(Leaf_Node)

  # ------------------------------------------------------------
  # PRECISION (TRAIN) – ONLY observations in top-K leaves
  # ------------------------------------------------------------
  train_flagged_idx <- train_tmp$Leaf_Node %in% top_k_rules
  train_flagged     <- sum(train_flagged_idx)

  train_tp <- sum(
    train_flagged_idx &
    abs(train_tmp$.resid) >= top_5_percent_value
  )

  train_fp <- train_flagged - train_tp

  train_precision <- ifelse(
    train_flagged > 0,
    train_tp / train_flagged,
    NA_real_
  )

  # ------------------------------------------------------------
  # PRECISION (TEST) – ONLY observations in top-K leaves
  # ------------------------------------------------------------
  test_tmp <- bind_cols(
    test_set,
    Leaf_Node = predict(model, test_set)
  )

  test_flagged_idx <- test_tmp$Leaf_Node %in% top_k_rules
  test_flagged     <- sum(test_flagged_idx)

  test_tp <- sum(
    test_flagged_idx &
    abs(test_tmp$.resid) >= top_5_percent_value
  )

  test_fp <- test_flagged - test_tp

  test_precision <- ifelse(
    test_flagged > 0,
    test_tp / test_flagged,
    NA_real_
  )

  # ------------------------------------------------------------
  # SAVE RUN
  # ------------------------------------------------------------
  sensitivity_results <- bind_rows(
    sensitivity_results,
    tibble(
      run_id          = i,
      cp              = params$cp,
      minbucket       = params$minbucket,
      maxdepth        = params$maxdepth,
      vars            = params$vars,
      train_lift      = train_lift,
      test_lift       = test_lift,
      full_lift       = full_lift,
      train_precision = train_precision,
      test_precision  = test_precision,
      train_flagged   = train_flagged,
      test_flagged    = test_flagged,
      train_tp        = train_tp,
      train_fp        = train_fp,
      test_tp         = test_tp,
      test_fp         = test_fp
    )
  )
}



# ------------------------------------------------------------------
# 4. Select best model (by train lift)
# ------------------------------------------------------------------
best_params <- sensitivity_results %>%
  filter(!is.na(train_lift)) %>%
  arrange(desc(train_lift)) %>%
  slice(1)

cat("\n=============================\n")
cat("  BEST GRID SEARCH MODEL\n")
cat("=============================\n")
print(best_params, width = Inf)

 View(sensitivity_results)
 write.csv(sensitivity_results, "treealert_sensitivity_analysis2.csv", row.names = FALSE)

```
Train Final Model using Best Parameters
```{r}
###############################################################################
## Train Final Model using Best Parameters + Best Variable Combination
###############################################################################

# Build formula dynamically from best_params$vars
#final_formula <- as.formula(paste(".resid ~", best_params$vars))

cat("\n=============================\n")
cat("  FINAL MODEL FORMULA USED\n")
cat("=============================\n")
print(final_formula)

final_formula <- .resid ~  Hour + Weekday + Month 

# Train final rpart model
best_tree_model <- rpart(
  final_formula,
  data = train_set,
  method = "anova",
  control = rpart.control(
    cp        = 0.0001, #best_params$cp,
    minbucket = 5, #best_params$minbucket,
    maxdepth  = 15 #best_params$maxdepth
  )
)

# Display tree rules
#cat("\n=============================\n")
#cat("        FINAL TREE RULES\n")
#cat("=============================\n")
#tree_rules <- rpart.rules(best_tree_model, cover = FALSE, digits = 4, 
#                          nn = FALSE)
#print(tree_rules)

# Attach predicted leaf node averages to full dataset
leaf_nodes <- predict(best_tree_model, newdata = data_df, type = "vector")
heart_merged_with_leaf <- dplyr::bind_cols(data_df, Leaf_Node = leaf_nodes)

# Preview result
cat("\n=============================\n")
cat(" ATTACHED LEAF NODE PREDICTIONS\n")
cat("=============================\n")
#print(head(heart_merged_with_leaf))
```
## LIFT RESULTS FOR BEST GRID SEARCH MODEL (Train / Test / Full)
```{r}
# ---- Train lift (Best model) ----
best_train_lift <- compute_lift_numeric(
  df         = train_set %>%
               dplyr::bind_cols(
                 Leaf_Node = predict(best_tree_model, newdata = train_set)
               ),
  actual_col = ".resid",
  pred_col   = "Leaf_Node",
  n_groups   = 20
)

# ---- Test lift (Best model) ----
best_test_lift <- compute_lift_numeric(
  df         = test_set %>%
               dplyr::bind_cols(
                 Leaf_Node = predict(best_tree_model, newdata = test_set)
               ),
  actual_col = ".resid",
  pred_col   = "Leaf_Node",
  n_groups   = 20
)

# ---- Full-sample lift (Best model) ----
best_full_lift <- compute_lift_numeric(
  df         = heart_merged_with_leaf,
  actual_col = ".resid",
  pred_col   = "Leaf_Node",
  n_groups   = 20
)

# ---- Report results ----
cat("\n=============================\n")
cat(" BEST GRID SEARCH MODEL LIFT\n")
cat("=============================\n")
cat("Train Lift (First Ventile):", best_train_lift, "\n")
cat("Test  Lift (First Ventile):", best_test_lift, "\n")
#cat("Full  Lift (First Ventile):", best_full_lift, "\n")
```
# -- DATA VIZ --
```{r, warning=FALSE, message=FALSE}
#Map rules to observations and  pull top 10 rules
start_row <- 1
end_row   <- TOP_K_RULES

# ------------Pull top K rules by leaf node value------------
leaf_summary <- heart_merged_with_leaf %>%
  filter(Time >= as.POSIXct(train_start, tz = "UTC") &
         Time <  as.POSIXct(train_end,   tz = "UTC")) %>%
  group_by(Leaf_Node) %>%
  summarise(mean_resid = mean(.resid, na.rm = TRUE), .groups = "drop") %>%
  arrange(desc(abs(mean_resid)))

top_k_leaf_values <- leaf_summary %>%
  slice(1:TOP_K_RULES) %>%
  pull(Leaf_Node)

#applies rules as filter and plot results

#------------Filter the rows for each rule for plotting------------
list_of_dfs <- list() # Create an empty DataFrame to store the results
for (i in start_row:end_row) {

  filtered_data <- heart_merged_with_leaf %>%
    filter(Leaf_Node == top_k_leaf_values[i])
    
  #------------displays the rule text before each plot------------
  # Pull the first row from tree_rules
  first_row <- tree_rules[i, ]
  # Filter out blank cells
  non_blank_cells <- first_row[first_row != ""]
  # Convert the non-blank cells to a single string using paste
  constructed_sentence <- paste(names(non_blank_cells), non_blank_cells, collapse = " ")
  rule_text <- textGrob(paste("Rule", i, "\n", constructed_sentence),
                      gp = gpar(col = "black", fontsize = 9))

    print(constructed_sentence)
    
  #------------Plot of filtered rules on fitted time scale------------
  post_rules_scatter_plot <- ggplot(filtered_data, 
                                    aes(x = Time, y = .resid)) + 
    geom_point(aes(color = .resid > 10)) +
    geom_text(aes(label = format(Time, "%H:%M")), 
              hjust = -0.1, vjust = 0.5, size = 2) +  # Adding time labels
    labs(title = "Scatter Plot Highlighted of Residuals", x = "Time", y = "Residual") +
      scale_color_manual(values = c("FALSE" = "#FF0000", "TRUE" = "#9400D3")) +
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 2)) 

  #------------Create Residual Plot------------
  #Initiate arguments
  #shape_number <- ifelse(row <= 25, row, 20)
  random_color <- rgb(runif(1), runif(1), runif(1), maxColorValue = 255)
  list_of_dfs[[i]] <- filtered_data # used to save df's for lower plot.
  
  # Plot of filtered rules on full series
  post_rules_line_plot <- ggplot(data_df, 
                                 aes(x = Time, y = .resid)) +
    geom_line(color = "#00BFC4") +
    geom_point(data = filtered_data, 
               aes(x = Time, y = .resid,
                   color = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
                   shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
                   size = 2.5, stroke = 0.8) +
    scale_color_manual(values = c("FALSE" = color_string[i], 
                                  "TRUE" = color_string[i])) +  
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
    labs(title = paste("Residual Plot with Tree Filtered Extreme and Non-extreme Values",
                       sep=""), x = NULL, y = NULL) +
    theme_bw() +
    theme(legend.position = "none") + 
    theme(plot.title = element_markdown())

    #------------Plot of Actual Vs Predicted------------
  actual_predicted_plot <- ggplot(data_df, aes(x = Time)) +
    geom_line(aes(y = Value, color = "Actual"), show.legend = FALSE) +
    geom_point(data = filtered_data, 
               aes(x = Time, y = Value, 
      shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
      color = .resid >= top_5_percent_value | .resid <=bottom_5_percent_value),
                   size = 2.5, stroke = 1, show.legend = FALSE) +  
    scale_color_manual(values = c("TRUE" = color_string[i], 
                                  "FALSE" = color_string[i], 
                                  "Actual" = "#4682B4"),
                       guide = FALSE) + 
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +  
    labs(title = "<span style='color:#4682B4'><b>Actual</b>
         </span> Electricity Demand + Tree Filtered Points",
         x = NULL, y = NULL, color = "Variable") +
    theme(legend.position = "none") +
    theme_bw() +
    theme(plot.title = element_markdown()) 

    #Output plots
     grid.arrange(rule_text, post_rules_scatter_plot, ncol = 1)
     grid.arrange(post_rules_line_plot, actual_predicted_plot, 
                  ncol = 1, heights = c(10,10))
}
```
All Rules Plots
```{r, warning=FALSE, message=FALSE}
first_test_date <- min(test_set$Time) # marks the date of split

#------------Filtered Actual Vs Predicted Plot------------
actual_predicted_plot <- ggplot(data_df, aes(x = Time)) +
  geom_line(aes(y = Value, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  geom_line(aes(y = .fitted, color = "Fitted"), 
            alpha = 0.6, show.legend = FALSE) +
  add_fitted_points_func(list_of_dfs, color_string, 
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
  scale_color_manual(values = c("Actual" = "#4682B4", "Fitted" = "#FF8C00")) +
  labs(title = "<span style='color:#4682B4'><b>Actual</b></span> & 
       <span style='color:#FF8C00'><b>Fitted</b></span> Electricity Demand", 
       x = NULL, y = NULL, color = "Variable") +
  geom_vline(xintercept = as.numeric(first_test_date), linetype = "dotted", color = "red", size = 1) +  # Add red dotted line
  theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10))
actual_predicted_plot

#------------Filtered Residual Plots------------
residual_plot <- ggplot(data_df , aes(x = Time)) +
  geom_line(aes(y = .resid, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  add_resid_points_func(list_of_dfs, color_string, # Generate geom_point layers
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
  scale_color_manual(values = c("Actual" = "#00BFC4")) +
  labs(title = "CRPS", x = NULL, y = NULL, color = "Variable") + 
  geom_vline(xintercept = as.numeric(first_test_date), linetype = "dotted", 
             color = "red", size = 1) +  # Add red dotted line
  theme_bw() + theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10)) 
residual_plot

two_plots <- grid.arrange(actual_predicted_plot, residual_plot, ncol = 1)
# Export the plot
ggsave("Electricity4.png", 
       plot = two_plots, width = 6, height = 4.5, dpi = 800) 
```

Calculate Precision and Related Metrics 
```{r}
#Calculate Precision for Train Set
train_filtered <- heart_merged_with_leaf %>%
  filter(Time >= as.POSIXct(train_start, tz = "UTC") & 
         Time <  as.POSIXct(train_end,   tz = "UTC"))

flagged_times <- bind_rows(list_of_dfs) %>%
  filter(Time >= as.POSIXct(train_start, tz = "UTC") & 
         Time <  as.POSIXct(train_end,   tz = "UTC")) %>%
  pull(Time)

train_metrics <- train_filtered %>%
  mutate(
    is_actual_extreme = (.resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
    is_flagged = Time %in% flagged_times,

    true_positive  = is_actual_extreme & is_flagged,
    false_positive = !is_actual_extreme & is_flagged
  )

tp <- sum(train_metrics$true_positive)
fp <- sum(train_metrics$false_positive)

precision <- tp / (tp + fp)

cat("\n--- Train Set Performance ---\n")
cat("Total flagged points:", tp + fp, "\n")
cat("Correctly identified extremes (TP):", tp, "\n")
cat("Incorrectly flagged (FP):", fp, "\n")
cat("Precision:", round(precision * 100, 2), "%\n")

# Calculate Precision for Test Set

test_filtered <- heart_merged_with_leaf %>%
  filter(Time >= as.POSIXct(test_start, tz = "UTC") & 
         Time <  as.POSIXct(test_end,   tz = "UTC"))

flagged_times_test <- bind_rows(list_of_dfs) %>%
  filter(Time >= as.POSIXct(test_start, tz = "UTC") & 
         Time <  as.POSIXct(test_end,   tz = "UTC")) %>%
  pull(Time)

test_metrics <- test_filtered %>%
  mutate(
    is_actual_extreme = (.resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
    is_flagged = Time %in% flagged_times_test,

    true_positive  = is_actual_extreme & is_flagged,
    false_positive = !is_actual_extreme & is_flagged
  )

tp_test <- sum(test_metrics$true_positive)
fp_test <- sum(test_metrics$false_positive)

precision_test <- tp_test / (tp_test + fp_test)

cat("\n--- Test Set Performance ---\n")
cat("Total flagged points:", tp_test + fp_test, "\n")
cat("Correctly identified extremes (TP):", tp_test, "\n")
cat("Incorrectly flagged (FP):", fp_test, "\n")
cat("Precision:", round(precision_test * 100, 2), "%\n")
```
# TreeAlert Sensitivity Analysis – Visualization Block
```{r}
library(tidyverse)
library(ggplot2)
library(scales)
library(corrplot)

# ------------------------------------------------------------
# Load sensitivity results
# ------------------------------------------------------------
sens <- sensitivity_results

# ------------------------------------------------------------
# 1. Lift–Precision Tradeoff Scatter Plot
# ------------------------------------------------------------
ggplot(
  sens,
  aes(
    x = test_lift,
    y = test_precision,
    color = factor(maxdepth),
    size = test_flagged
  )
) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(
    name = "Flagged periods (test)",
    labels = comma
  ) +
  scale_color_brewer(
    palette = "Set1",
    name = "Tree max depth"
  ) +
  labs(
    title = "Lift–Precision Tradeoff Across TreeAlert Configurations",
    x = "Test Lift (First Ventile)",
    y = "Test Precision"
  ) +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# ------------------------------------------------------------
# 2. Stability Plot Across Variable Sets (Mean ± SD)
# ------------------------------------------------------------
vars_stability <- sens %>%
  group_by(vars) %>%
  summarise(
    mean_test_lift = mean(test_lift, na.rm = TRUE),
    sd_test_lift   = sd(test_lift, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_test_lift))

ggplot(
  vars_stability,
  aes(
    x = reorder(vars, mean_test_lift),
    y = mean_test_lift
  )
) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(
      ymin = mean_test_lift - sd_test_lift,
      ymax = mean_test_lift + sd_test_lift
    ),
    width = 0.2
  ) +
  coord_flip() +
  labs(
    title = "Stability of TreeAlert Performance Across Variable Sets",
    x = "Temporal variable combination",
    y = "Mean test lift ± SD"
  ) +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# ------------------------------------------------------------
# 3. Flagged Volume vs Performance (Operational Tradeoff)
# ------------------------------------------------------------
ggplot(
  sens,
  aes(
    x = test_flagged,
    y = test_lift,
    color = factor(minbucket)
  )
) +
  geom_point(alpha = 0.7) +
  scale_x_continuous(labels = comma) +
  scale_color_brewer(
    palette = "Dark2",
    name = "Min bucket"
  ) +
  labs(
    title = "Operational Tradeoff: Alert Volume vs Test Performance",
    x = "Flagged periods (test)",
    y = "Test lift (first ventile)"
  ) +
  theme_bw() +
  theme(plot.title = element_text(face = "bold"))

# ------------------------------------------------------------
# 4. Correlation Matrix of Sensitivity Variables
# ------------------------------------------------------------
num_vars <- sens %>%
  select(
    cp, minbucket, maxdepth,
    train_lift, test_lift, full_lift,
    train_precision, test_precision,
    train_flagged, test_flagged
  )

corr_mat <- cor(num_vars, use = "pairwise.complete.obs")

corrplot(
  corr_mat,
  method = "color",
  type = "upper",
  tl.col = "black",
  tl.cex = 0.8,
  addCoef.col = "black",
  number.cex = 0.6
)

```

One time Pre-processing - old stuff. Ignore
```{r, include=FALSE, warning=FALSE, message=FALSE}
# test_data <- read.csv(
#   "C:/Users/Kagebi/Documents/Research/TreeAlert and Lift/Power_Lift/ensemble_metrics_timeseries_test.csv"
# )
# 
# train_data <- read.csv(
#   "C:/Users/Kagebi/Documents/Research/TreeAlert and Lift/Power_Lift/train_aggregation_metrics_timeseries.csv"
# )

# # fix sci notation

# avoid scientific notation
# options(scipen = 999)

# function to round quantiles + winkler
# round_metrics <- function(df, digits = 4) {
#   q_cols <- grep("^q", names(df), value = TRUE)
#   cols_to_round <- c(q_cols, "winkler_score")
# 
#   df[cols_to_round] <- lapply(
#     df[cols_to_round],
#     function(x) round(as.numeric(x), digits)
#   )
#   df
# }

# apply to train and test
# train_data <- round_metrics(train_data, digits = 4)
# test_data  <- round_metrics(test_data,  digits = 4)
## Calc CRPS
# library(scoringRules)
# 
# q_cols <- grep("^q", names(test_data), value = TRUE)
# 
# # Calculate CRPS for test_data
# test_data$crps <- sapply(1:nrow(test_data), function(i) {
#   quantiles <- as.numeric(test_data[i, q_cols])
#   actual <- test_data$actual[i]
#   crps_sample(actual, quantiles)
# })
# 
# # Calculate CRPS for train_data
# train_data$crps <- sapply(1:nrow(train_data), function(i) {
#   quantiles <- as.numeric(train_data[i, q_cols])
#   actual <- train_data$actual[i]
#   crps_sample(actual, quantiles)
# })
# 
# # Round to 4 decimal places
# test_data$crps <- round(test_data$crps, 4)
# train_data$crps <- round(train_data$crps, 4)

##Save as RData files
# save(train_data, file = "train_r_v2.RData")
# save(test_data, file = "test_r_v2.RData")

# library(data.table)
# 
# # Use normalizePath to handle Windows backslash issues automatically
# base_path <- "C:/Users/Kagebi/Documents/Research/TreeAlert and Lift/solar_wind/"
# 
# # Load files explicitly
# load(file.path(base_path, "train_r_v2.RData"), verbose = TRUE)
# load(file.path(base_path, "test_r_v2.RData"), verbose = TRUE)
# 
# # remove extra data in train and save as 2 train files
# 
# 
# # 1. Convert train_data to a data.table in-place
# setDT(train_data)
# 
# # 2. Convert valid_datetime to POSIXct for accurate date comparison
# # This handles the "YYYY-MM-DD HH:MM:SS" format seen in your images
# train_data[, valid_datetime := as.POSIXct(valid_datetime, format = "%Y-%m-%d %H:%M:%S")]
# 
# # 3. Define the cutoff and filter
# # Removes all rows >= 2/1/2023 0:00
# cutoff_date <- as.POSIXct("2023-02-01 00:00:00")
# train_data <- train_data[valid_datetime < cutoff_date]
# 
# # 4.1 Split based on the 'type' column
# train_wind  <- train_data[type == "wind"]
# train_solar <- train_data[type == "solar"]
# 
# # 5. Define your path and save as separate RData files
# path <- "C:/Users/Kagebi/Documents/Research/TreeAlert and Lift/solar_wind/"
# 
# save(train_wind, file = paste0(path, "train_wind_v2.RData"))
# save(train_solar, file = paste0(path, "train_solar_v2.RData"))
# 
# # Quick verification
# print(nrow(train_wind))
# print(nrow(train_solar))
# 
# # Split the test data
# 
# setDT(test_data)
# # 4.2 Split based on the 'type' column
# test_wind  <- test_data[type == "wind"]
# test_solar <- test_data[type == "solar"]
# 
# save(test_wind, file = paste0(path, "test_wind_v2.RData"))
# save(test_solar, file = paste0(path, "test_solar_v2.RData"))

# 
# ## new data 
# library(data.table)
# 
# # 1. Ensure both are data.tables
# setDT(train_wind)
# setDT(test_wind)
# 
# # 2. Rename 'case' to 'split' in test_wind if it hasn't been done yet
# if ("case" %in% names(test_wind)) {
#   setnames(test_wind, "case", "split")
# }
# 
# # 3. Standardize valid_datetime
# # Convert test_wind (character) to POSIXct to match train_wind (POSIXct)
# if (is.character(test_wind$valid_datetime)) {
#   test_wind[, valid_datetime := as.POSIXct(valid_datetime, format = "%m/%d/%Y %H:%M")]
# }
# 
# # 4. Remove scientific notation and round numeric columns
# options(scipen = 999)
# num_cols <- names(train_wind)[sapply(train_wind, is.numeric)]
# 
# train_wind[, (num_cols) := lapply(.SD, function(x) round(x, 4)), .SDcols = num_cols]
# test_wind[, (num_cols) := lapply(.SD, function(x) round(x, 4)), .SDcols = num_cols]
# 
# # 5. Merge - stacking test at the end of train
# all_wind_data <- rbind(train_wind, test_wind, use.names = TRUE)
# 
# # 6. Save to raw_data folder
# save_path <- "C:/Users/Kagebi/Documents/Research/TreeAlert and Lift/raw_data/"
# if (!dir.exists(save_path)) dir.create(save_path, recursive = TRUE)
# 
# save(all_wind_data, file = file.path(save_path, "all_wind_merged.RData"))
# 
# # 7. Final Verification
# print("Merge complete. Summary of valid_datetime:")
# #summary(all_wind_data$valid_datetime)
# 
# ##  Create the Mean Forecast Column
# # 1. Ensure it's a data.table
# setDT(all_wind_data)
# 
# # 2. Create the 'forecast' column directly (Mean of all columns starting with 'q')
# # We use .SDcols to select all quantile columns dynamically
# q_cols <- grep("^q", names(all_wind_data), value = TRUE)
# all_wind_data[, forecast := rowMeans(.SD, na.rm = TRUE), .SDcols = q_cols]
```


