---
title: "Eval Tree Analysis Code"
author: "Bobea"
date: "2024-02-03"
output: html_document
---
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      TreeAlert Electricity Example                         
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
Libraries
```{r, include=FALSE, warning=FALSE, message=FALSE}
library(readr); library(tidyr); library(feasts) # For Data
library(tsibble); library(fable); library(fabletools) # For Forecasting
library(dplyr); library(lubridate) # For Data Wrangling
library(ggplot2); library(ggtext); library(grid); library(gridExtra); # Plot
library(slider) # Plot
library(scales); library(rpart);  library(rpart.plot);
```
Define Functions:
```{r}
# temporal_encoding <- function(data) {
#   data <- data %>%
#     mutate(
#       Time = parse_date_time(Time, orders = c("ymd HMS", "ymd HM", "ymd H", "ymd")),  # Try multiple formats
#       Hour = factor(hour(Time)),
#       Minute = factor(minute(Time)),
#       Weekday = factor(wday(Time, label = TRUE)),
#       DayofMonth = factor(mday(Time)),
#       WeekofMonth = factor(ceiling(day(Time) / 7)),
#       Month = factor(month(Time, label = TRUE))
#     ) %>%
  # return(data)
  #   }
temporal_encoding <- function(data) {
  # Convert Time column to Date-Time format and create temporal features
   data <- data %>%
    mutate(
      Time = parse_date_time(Time, orders = c("ymd HMS", "ymd HM", "ymd H", "ymd")),
      Hour = factor(hour(Time)),
      Minute = factor(minute(Time)),
      Weekday = factor(lubridate::wday(Time, label = TRUE)),
      DayofMonth = factor(mday(Time)),
      WeekofMonth = factor(ceiling(day(Time) / 7)),
      Month = factor(lubridate::month(Time, label = TRUE, abbr = TRUE))
    ) %>%
    arrange(Time)

  # Add lagged residuals for 1–4 weeks
  for (week in 1:4) {
    lag_label <- paste0("Lag_", week, "_Week")
    lag_time_col <- paste0("LagTime_", week, "_Week")

    data <- data %>%
      mutate(!!lag_time_col := Time - weeks(week)) %>%
      left_join(
        data %>%
          select(Time, .resid) %>%
          rename(!!lag_label := .resid),
        by = setNames("Time", lag_time_col)
      ) %>%
      select(-all_of(lag_time_col))
  }

  # Fill NAs in all lag columns with previous value
  data <- data %>%
    arrange(Time) %>%
    fill(starts_with("Lag_"), .direction = "down")

  return(data)
}

# process_data converts a random df into a format usable by Tree Alert!
process_data <- function(data, time_col, actual_col, forecast_col) {

  data %>%
    # Select specified columns and rename them
    dplyr::select({{ time_col }}, {{ actual_col }}, {{ forecast_col }}) %>%
    dplyr::rename(
      Time = {{ time_col }},         # Rename timestamp
      actual = {{ actual_col }},     # Rename actual load
      forecast = {{ forecast_col }}  # Rename forecast load
    ) %>%
    # Convert to datetime and handle multiple possible formats
    dplyr::mutate(
      Time = parse_date_time(Time, orders = c("ymd HMS", "ymd HM", "mdy HMS", "mdy HM", "dmy HMS", "dmy HM")),
      .resid = actual - forecast     # Calculate forecast error
    ) %>%
    na.omit()                        # Remove any NA values
}

# To generate geom_point layers for 'All Filters Actual Vs Predicted'
add_fitted_points_func <- function(data_list, color_list, shape_values) {
  layers <- lapply(seq_along(data_list), function(i) {
    geom_point(data = data_list[[i]], 
               aes(x = Time, y = Value, shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
               size = 2.5, stroke = 1, color = color_list[i])
  })
  return(layers)
}

# To generate geom_point layers for 'All Filters Residuals'
add_resid_points_func <- function(data_list, color_list, shape_values) {
  layers <- lapply(seq_along(data_list), function(i) {
    geom_point(data = data_list[[i]], 
               aes(x = Time, y = .resid, shape = .resid >= top_5_percent_value | 
                     .resid <= bottom_5_percent_value),
               size = 2.5, stroke = 1, color = color_list[i])
  })
  return(layers)
}

# Function to create the plot for a specific month
create_monthly_plot <- function(data, start_date, end_date) {
  plot <- ggplot(data, aes(x = Time)) +
    geom_line(aes(y = Value, color = "Actual"), show.legend = FALSE) +
    geom_line(aes(y = model_1, color = "Fitted"), alpha = 0.6, show.legend = FALSE) +
    # changed to skip rules 1-2 
    add_fitted_points_func(list_of_dfs[start_row:length(list_of_dfs)], 
                       color_string[start_row:length(list_of_dfs)], 
                       c("TRUE" = 1, "FALSE" = 16))
    #add_fitted_points_func(list_of_dfs, color_string, # make geom_point layers
    #                                  c("TRUE" = 1, "FALSE" = 16)) + 
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
    scale_color_manual(values = c("Actual" = "#4682B4", "Fitted" = "#FF8C00")) +
    scale_x_datetime(limits = as.POSIXct(c(start_date, end_date)), 
                     date_labels = "%m-%d",
                     #date_breaks = "3 days",
                     expand = c(0.02, 0.02)) +
    labs(title = paste(""), x = NULL, y = NULL, color = "Variable") +
    theme_bw() +
    theme(legend.position = "none",
          plot.title = element_markdown(size = 10),
          plot.title.position = "panel")
  
  return(plot)
}

# -------------------- Lift Functions --------------------
compute_lift_numeric <- function(df, actual_col, pred_col, n_groups = 33) {
    
  actual <- abs(df[[actual_col]])
  score  <- abs(df[[pred_col]])
  
  # Sort by predicted risk
  ord        <- order(score, decreasing = TRUE)
  actual_ord <- actual[ord]
  n          <- length(actual_ord)
  
  # Ventile grouping
  group <- ceiling(seq_len(n) / (n / n_groups))
  group[group > n_groups] <- n_groups
  
  df_lift <- data.frame(group = group, error = actual_ord) %>%
    dplyr::group_by(group) %>%
    dplyr::summarise(meanError = mean(error, na.rm = TRUE), .groups = "drop")
  
  # Normalize
  df_lift$meanError <- df_lift$meanError / mean(actual_ord, na.rm = TRUE)
  
  # Return first ventile lift
  df_lift$meanError[df_lift$group == 1]
}

make_lift_charts <- function(df, actual_col, pred_col, title_prefix = "Model") {
  
  actual <- abs(df[[actual_col]])
  score  <- abs(df[[pred_col]])
  
  # Sort by predicted risk
  ord <- order(score, decreasing = TRUE)
  actual_ord <- actual[ord]
  n <- length(actual_ord)
  
  # Cumulative gains
  df_cum <- data.frame(
    ncases  = c(0, seq_len(n)),
    cumError = c(0, cumsum(actual_ord))
  )
  
  # quantiles
  n_groups <- 50
  group <- ceiling(seq_len(n) / (n / n_groups))
  group[group > n_groups] <- n_groups
  
  df_lift <- data.frame(
    group = group,
    error = actual_ord
  ) %>%
    dplyr::group_by(group) %>%
    dplyr::summarise(meanError = mean(error), .groups = "drop")
  
  df_lift$percentile <- df_lift$group * 5
  df_lift$meanError  <- df_lift$meanError / mean(actual_ord)
  
  g1 <- ggplot(df_cum, aes(ncases, cumError)) +
    geom_line(color = "#4682B4") +
    geom_abline(slope = sum(actual_ord)/n, intercept = 0,
                linetype = "dashed", color = "gray") +
    labs(title = paste(title_prefix, "– Cumulative Gains"),
         x = "# Cases", y = "Cumulative Error") +
    theme_minimal()
  
  g2 <- ggplot(df_lift, aes(percentile, meanError)) +
    geom_col(fill = "#FF8C00") +
    labs(title = paste(title_prefix, "– Ventile Lift"),
         x = "Percentile", y = "Ventile Mean / Global Mean") +
    theme_minimal()
  
  gridExtra::grid.arrange(g1, g2, ncol = 2)
  
  df_lift$meanError[df_lift$percentile == 5]
}


# Prep Parameters
color_string <- c("#9400D3", "#008000", "#0000FF", "#FF8C00", "#000066",
                  "#FF00FF", "#AB0000", "#FF69B4", "#7B68EE", "#Ff0000", 
                  "#7B68EE", "#20B2AA", "#703366", "#32CD32", "#CC6633", 
                  "#C77CFF", "#8A2BE2", "#00FF7F", "#00BFFF", "#FFA500", 
                  "#66CDAA", "#BA55D3", "#00CED1", "#4682B4", "#00FFFF")
```
Load files from Module A
```{r, warning=FALSE}
raw_data <- read.csv("C:/Users/Kagebi/Documents/Research/time_series_15min_singleindex.csv") %>%
  slice(-(1:5))
```
# START!  
Select your data!
```{r}
# Region,Actual Load,Forecasted Load
# Austria (AT): AT_load_actual_entsoe_transparency, AT_load_forecast_entsoe_transparency
# Belgium (BE): BE_load_actual_entsoe_transparency,BE_load_forecast_entsoe_transparency
# Germany: DE_load_actual_entsoe_transparency,DE_load_forecast_entsoe_transparency
# Germany-Lux (DE-LU BZ): DE_LU_load_actual_entsoe_transparency,DE_LU_load_forecast_entsoe_transparency
# DE Control Area 50Hertz: DE_50hertz_load_actual_entsoe_transparency,DE_50hertz_load_forecast_entsoe_transparency
# DE Control Area Amprion: DE_amprion_load_actual_entsoe_transparency,DE_amprion_load_forecast_entsoe_transparency
# DE Control Area TenneT GER: DE_tennet_load_actual_entsoe_transparency,DE_tennet_load_forecast_entsoe_transparency
# DE Control Area TransnetBW: DE_transnetbw_load_actual_entsoe_transparency,DE_transnetbw_load_forecast_entsoe_transparency
# Hungary (HU): HU_load_actual_entsoe_transparency,HU_load_forecast_entsoe_transparency
# Luxembourg (LU): LU_load_actual_entsoe_transparency,LU_load_forecast_entsoe_transparency
# Netherlands (NL): NL_load_actual_entsoe_transparency,NL_load_forecast_entsoe_transparency

# Define the specific column names you want to keep
Actual_column <- "DE_LU_load_actual_entsoe_transparency"
Forecast_column <- "DE_LU_load_forecast_entsoe_transparency"

data_start <-"2019-02-01" 
data_end <-"2019-5-30"

train_start <- "2019-02-01"
train_end <- "2019-5-01"

test_start <- train_end 
test_end<- data_end 

#--------------------------------------------------------
# Select only the timestamp columns (1-2) and your specified columns using dplyr
data_df <- raw_data %>%
  select(1:2, all_of(Actual_column), all_of(Forecast_column))

data_df <- process_data(
  raw_data, 
  time_col = utc_timestamp, 
  actual_col = Actual_column, 
  forecast_col = Forecast_column
)

# Reduce data size with a Time Filter (Optional)
data_df <- data_df[data_df$Time >= as.POSIXct(data_start, tz = "UTC"), ]
data_df <- data_df[data_df$Time < as.POSIXct(data_end, tz = "UTC"), ]

# old chart
# data_df <- data_df[data_df$Time >= as.POSIXct("2020-05-01", tz = "UTC"), ]
# data_df <- data_df[data_df$Time <= as.POSIXct("2020-08-1", tz = "UTC"), ]
  # actual_col = LU_load_actual_entsoe_transparency, 
  # forecast_col = LU_load_forecast_entsoe_transparency cp 0 min bucket 4

## Previous chart: 
# data_start <-"2019-04-01" #need buffer for lag
# data_end <-"2019-6-07"
# 
# train_start <- "2019-04-08"
# train_end <- "2019-6-01"
```
Create a new data frame that matches the structure of data_df
```{r}
# Aggregate by hour for more clear plots
data_filtered <- data_df %>%
  mutate(hour = format(Time, "%Y-%m-%d %H")) %>%
  group_by(hour) %>%
  filter(actual == max(actual)) %>%
  ungroup() %>%
  select(-hour) 

data_df <- temporal_encoding(data_filtered) 

# Create a copy with the required columns and modifications
data_df <- data_df %>%
  mutate(
    Value = actual,
    .fitted = forecast,
    .resid = .resid
  ) %>%
  select(-actual, -forecast) %>%
  as_tibble()  # Convert to tibble if consistency is needed

#------------Train Test ------------
  
# Determine the split index (80% of the data)
#split_index <- floor(0.9 * nrow(data_df))
# Split into training and testing sets
#train_set <- data_df[1:split_index, ]
#test_set <- data_df[(split_index + 1):nrow(data_df), ] #Not needed. Just quick for viewing

#option 2 --  use time to split
train_set <- data_df[data_df$Time > as.POSIXct(train_start, tz = "UTC") & 
                    data_df$Time < as.POSIXct(train_end, tz = "UTC"),]
test_set <- data_df[data_df$Time >= as.POSIXct(test_start, tz = "UTC") & 
                    data_df$Time < as.POSIXct(test_end, tz = "UTC"),]
# train_set <- train_set %>%
  # mutate(resid_log = log1p(abs(.resid)) * sign(.resid))  # preserves sign

#------------Top bottom 5% of data tree ------------
# Calculate quantiles
top_5_percent <- quantile(train_set$.resid, probs = 0.95, na.rm = TRUE)
bottom_5_percent <- quantile(train_set$.resid, probs = 0.05, na.rm = TRUE)
# Extract values
top_5_percent_value <- top_5_percent[[1]] 
bottom_5_percent_value <- bottom_5_percent[[1]]
```
# Make Trees & Pull Rules
```{r}
#------------Regression tree on .resid + temporal variables------------
tree_model <- rpart(.resid ~ Minute + Hour + Weekday,
                    data = train_set, method = "anova", 
                    cp = 0.0000, minbucket = 5)
# cp = 0.000, minbucket = 3
# cp = 0.00006, minbucket = 2
# + DayofMonth+ WeekofMonth, + Lag_2_Week + Minute +
tree_rules <- rpart.rules(tree_model, cover = FALSE, digits = 4, nn = FALSE)

#Attached the predicted tree nodes to DF and Maps rules/leafs to Observations
leaf_nodes <- predict(tree_model, newdata = data_df, type = "vector")
heart_merged_with_leaf <- dplyr::bind_cols(data_df, Leaf_Node = leaf_nodes)
```
Lift Analysis
```{r}
compute_lift <- function(data, model, pred_col_name, label, plot = TRUE, n_groups = 50) {

  # Remove existing prediction column if present
  data <- data %>% dplyr::select(-dplyr::any_of(pred_col_name))

  # Predict leaf-node values (force numeric)
  leaf_nodes <- as.numeric(predict(model, newdata = data, type = "vector"))

  # Attach predictions
  data_with_pred <- dplyr::bind_cols(
    data,
    !!pred_col_name := leaf_nodes
  )

  # Plot or numeric lift
  lift_value <- if (plot) {
    make_lift_charts(
      df           = data_with_pred,
      actual_col   = ".resid",
      pred_col     = pred_col_name,
      title_prefix = label
    )
  } else {
    compute_lift_numeric(
      df         = data_with_pred,
      actual_col = ".resid",
      pred_col   = pred_col_name,
      n_groups   = n_groups
    )
  }

  cat(label, "Lift (First Ventile):", lift_value, "\n")

  return(list(
    lift        = lift_value,
    merged_data = data_with_pred
  ))
}
train_results <- compute_lift(train_set, tree_model, "Leaf_Node", "Baseline (Train)", plot = TRUE)
test_results  <- compute_lift(test_set,  tree_model, "Leaf_Node", "Baseline (Test)",  plot = TRUE)
full_results  <- compute_lift(data_df,   tree_model, "Leaf_Node", "Baseline (Full)",  plot = TRUE)
```
# Grid Search: Joint Optimization of Tree Parameters + Predictor Variables  ##
```{r}
# 1. Candidate predictors
candidate_vars <- c(
  "Hour", "Minute", "Weekday" , "DayofMonth", "WeekofMonth"
)

# 2. All non-empty combinations of predictors
var_combinations <- unlist(
  lapply( 1:length(candidate_vars), function(k) 
    combn(candidate_vars, k, FUN = paste, collapse = " + "))
)

# 3. Parameter grid INCLUDING predictors
param_grid <- expand.grid(
  cp        = c(0, 0.0001, 0.001, 0.005, 0.01),
  minbucket = c(5, 10, 20, 30),
  maxdepth  = c(5, 10, 15),
  vars      = var_combinations,
  stringsAsFactors = FALSE
)

# ----------------------------------------------------------------------------
# MODEL EVALUATION FUNCTION / Uses compute_lift_numeric() exactly as your main lift code
#Grid evaluation function
evaluate_params <- function(params_row) {

  cp        <- params_row$cp
  minbucket <- params_row$minbucket
  #maxdepth  <- params_row$maxdepth
  #vars      <- params_row$vars

  # Build formula dynamically
  fmla <- as.formula(paste(".resid ~", vars))

  # Train regression tree
  model <- rpart( fmla,
    data   = train_set,
    method = "anova",
    control = rpart.control(
      cp        = cp,
      minbucket = minbucket,
      maxdepth  = maxdepth
    )
  )

  # Predict leaf values on TRAIN SET ONLY
  train_pred <- predict(model, newdata = train_set, type = "vector")
 
   # Build temporary dataset to pass into compute_lift_numeric (df version)
  train_df <- dplyr::bind_cols(train_set, Leaf_Node = train_pred )

  # Compute lift (same function as everywhere else)
  lift_value <- compute_lift_numeric(
    df         = train_df,
    actual_col = ".resid",
    pred_col   = "Leaf_Node",
    n_groups   = 20
  )

  # Return results
  data.frame(
    cp                 = round(cp, 6),
    minbucket          = as.integer(minbucket),
    maxdepth           = as.integer(maxdepth),
    vars               = vars,
    first_ventile_lift = lift_value
  )
}

results <- do.call(
  rbind,
  lapply(seq_len(nrow(param_grid)), function(i) {
    evaluate_params(param_grid[i, , drop = FALSE])
  })
)

# RUN GRID SEARCH
results <- do.call(
  rbind,
  lapply(seq_len(nrow(param_grid)), function(i) {
    evaluate_params(param_grid[i, , drop = FALSE])
  })
)

# SELECT BEST MODEL
best_params <- results %>%
  dplyr::filter(!is.na(first_ventile_lift)) %>%
  dplyr::arrange(dplyr::desc(first_ventile_lift)) %>%
  dplyr::slice(1)

cat("\n=============================\n")
cat("  BEST TREE PARAMETERS FOUND\n")
cat("=============================\n")
print(best_params)
```
# Train Final Model using Best Parameters + LIFT
```{r}
# Build formula dynamically from best_params$vars
final_formula <- as.formula(paste(".resid ~", best_params$vars))

cat("\n=============================\n")
cat("  FINAL MODEL FORMULA USED\n")
cat("=============================\n")
print(final_formula)

# Train final rpart model
best_tree_model <- rpart(final_formula, data = train_set, method = "anova",
  control = rpart.control(
    cp        = best_params$cp,
    minbucket = best_params$minbucket,
    maxdepth  = best_params$maxdepth
  )
)

# Display tree rules
cat("\n=============================\n")
cat("        FINAL TREE RULES\n")
cat("=============================\n")
tree_rules <- rpart.rules(best_tree_model, cover = FALSE, digits = 4, nn = FALSE)
#print(tree_rules)

# Attach predicted leaf node averages to full dataset
leaf_nodes <- predict(best_tree_model, newdata = data_df, type = "vector")
heart_merged_with_leaf <- dplyr::bind_cols(data_df, Leaf_Node = leaf_nodes)

# Preview result
cat("\n=============================\n")
cat(" ATTACHED LEAF NODE PREDICTIONS\n")
cat("=============================\n")
print(head(heart_merged_with_leaf))



## LIFT RESULTS FOR BEST GRID SEARCH MODEL (Train / Test / Full)

# ---- Train lift (Best model) ----
best_train_lift <- compute_lift_numeric(
  df         = train_set %>%
               dplyr::bind_cols(
                 Leaf_Node = predict(best_tree_model, newdata = train_set)
               ),
  actual_col = ".resid",
  pred_col   = "Leaf_Node",
  n_groups   = 20
)

# ---- Test lift (Best model) ----
best_test_lift <- compute_lift_numeric(
  df         = test_set %>%
               dplyr::bind_cols(
                 Leaf_Node = predict(best_tree_model, newdata = test_set)
               ),
  actual_col = ".resid",
  pred_col   = "Leaf_Node",
  n_groups   = 20
)

# ---- Full-sample lift (Best model) ----
best_full_lift <- compute_lift_numeric(
  df         = heart_merged_with_leaf,
  actual_col = ".resid",
  pred_col   = "Leaf_Node",
  n_groups   = 20
)

# ---- Report results ----
cat("\n=============================\n")
cat(" BEST GRID SEARCH MODEL LIFT\n")
cat("=============================\n")
cat("Train Lift (First Ventile):", best_train_lift, "\n")
cat("Test  Lift (First Ventile):", best_test_lift, "\n")
cat("Full  Lift (First Ventile):", best_full_lift, "\n")
```
Map rules to observations and  pull top 10 rules
```{r}
# How many rows of tree rules would you like to use?
start_row <- 1 # starting row of rule from top_resid_rules
end_row <- 4   # ending row (should be <= 20). if last chart does not # # Ensure it's a standard data frame

# ------------Pull top 10 rules------------
tree_rules$.resid <- as.numeric(tree_rules$.resid)
tree_rules <- tree_rules[order(abs(tree_rules$.resid),
                                      decreasing = TRUE), ]
# # ------------Pull top rules using tail method (95th percentile)------------
# tree_rules$.resid <- as.numeric(tree_rules$.resid)
# 
# # Add Leaf_Node predictions to train_set
# train_with_leaf <- train_set %>%
#   mutate(Leaf_Node = predict(tree_model, newdata = train_set, type = "vector"))
# 
# # Calculate 95th percentile for each unique leaf node value
# tail_scores <- train_with_leaf %>%
#   group_by(Leaf_Node) %>%
#   summarise(tail95 = quantile(abs(.resid), probs = 0.95, na.rm = TRUE), .groups = "drop")
# 
# # Add tail95 to tree_rules by matching .resid to Leaf_Node
# tree_rules$tail95 <- tail_scores$tail95[match(tree_rules$.resid, tail_scores$Leaf_Node)]
# 
# # Sort by tail95 instead of mean absolute residual
# tree_rules <- tree_rules[order(tree_rules$tail95, decreasing = TRUE, na.last = TRUE), ]
# 
# # Remove the helper column after sorting (use base R)
# tree_rules$tail95 <- NULL


top_10_resid_list <- as.list(tree_rules$.resid) # not top 10 now
last_column_index <- ncol(top_10_resid_list)
color_string <- c("#9400D3", "#008000", "#0000FF", "#FF8C00", "#000066",
                  "#FF00FF", "#AB0000", "#FF69B4", "#7B68EE", "#Ff0000", 
                  "#7B68EE", "#20B2AA", "#703366", "#32CD32", "#CC6633", 
                  "#C77CFF", "#8A2BE2", "#00FF7F", "#00BFFF", "#FFA500", 
                  "#66CDAA", "#BA55D3", "#00CED1", "#4682B4", "#00FFFF")
```

# THE PROGRAM -- appllies rules as filter and plot results
```{r, warning=FALSE, message=FALSE}

#------------Filter the rows for each rule for plotting------------
list_of_dfs <- list() # Create an empty DataFrame to store the results
for (i in start_row:end_row) {

  # Pull the next iterative resid to filter by form the top 10 list
  current_rule <- as.numeric(top_10_resid_list[[i]])
  # round both to common decimal as heart_merged_with_leaf has 1 more decimal place
  precision <- 2  
  # Filter heart_merged_with_leaf
  filtered_data <- heart_merged_with_leaf %>%
    filter(round(Leaf_Node, precision) == round(current_rule, precision))
    
  #------------displays the rule text before each plot------------
  # Pull the first row from tree_rules
  first_row <- tree_rules[i, ]
  # Filter out blank cells
  non_blank_cells <- first_row[first_row != ""]
  # Convert the non-blank cells to a single string using paste
  constructed_sentence <- paste(names(non_blank_cells), non_blank_cells, collapse = " ")
  rule_text <- textGrob(paste("Rule", i, "\n", constructed_sentence),
                      gp = gpar(col = "black", fontsize = 9))

    print(constructed_sentence)
    
  #------------Plot of filtered rules on fitted time scale------------
  post_rules_scatter_plot <- ggplot(filtered_data, 
                                    aes(x = Time, y = .resid)) + 
    geom_point(aes(color = .resid > 10)) +
    geom_text(aes(label = format(Time, "%H:%M")), 
              hjust = -0.1, vjust = 0.5, size = 2) +  # Adding time labels
    labs(title = "Scatter Plot Highlighted of Residuals", x = "Time", y = "Residual") +
      scale_color_manual(values = c("FALSE" = "#FF0000", "TRUE" = "#9400D3")) +
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 2)) 

  #------------Create Residual Plot------------
  #Initiate arguments
  #shape_number <- ifelse(row <= 25, row, 20)
  random_color <- rgb(runif(1), runif(1), runif(1), maxColorValue = 255)
  list_of_dfs[[i]] <- filtered_data # used to save df's for lower plot.
  
  # Plot of filtered rules on full series
  post_rules_line_plot <- ggplot(data_df, 
                                 aes(x = Time, y = .resid)) +
    geom_line(color = "#00BFC4") +
    geom_point(data = filtered_data, 
               aes(x = Time, y = .resid,
                   color = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
                   shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
                   size = 2.5, stroke = 0.8) +
    scale_color_manual(values = c("FALSE" = color_string[i], 
                                  "TRUE" = color_string[i])) +  
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
    labs(title = paste("Residual Plot with Tree Filtered Extreme and Non-extreme Values",
                       sep=""), x = NULL, y = NULL) +
    theme_bw() +
    theme(legend.position = "none") + 
    theme(plot.title = element_markdown())

    #------------Plot of Actual Vs Predicted------------
  actual_predicted_plot <- ggplot(data_df, aes(x = Time)) +
    geom_line(aes(y = Value, color = "Actual"), show.legend = FALSE) +
    geom_point(data = filtered_data, 
               aes(x = Time, y = Value, 
      shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
      color = .resid >= top_5_percent_value | .resid <=bottom_5_percent_value),
                   size = 2.5, stroke = 1, show.legend = FALSE) +  
    scale_color_manual(values = c("TRUE" = color_string[i], 
                                  "FALSE" = color_string[i], 
                                  "Actual" = "#4682B4"),
                       guide = FALSE) + 
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +  
    labs(title = "<span style='color:#4682B4'><b>Actual</b>
         </span> Electricity Demand + Tree Filtered Points",
         x = NULL, y = NULL, color = "Variable") +
    theme(legend.position = "none") +
    theme_bw() +
    theme(plot.title = element_markdown()) 

    #Output plots
     grid.arrange(rule_text, post_rules_scatter_plot, ncol = 1)
     grid.arrange(post_rules_line_plot, actual_predicted_plot, 
                  ncol = 1, heights = c(10,10))
}
```
All Rules Plots
```{r, warning=FALSE, message=FALSE}
first_test_date <- min(test_set$Time) # marks the date of split

#------------Filtered Actual Vs Predicted Plot------------
actual_predicted_plot <- ggplot(data_df, aes(x = Time)) +
  geom_line(aes(y = Value, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  geom_line(aes(y = .fitted, color = "Fitted"), 
            alpha = 0.6, show.legend = FALSE) +
  add_fitted_points_func(list_of_dfs, color_string, 
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 1)) +
  scale_color_manual(values = c("Actual" = "#4682B4", "Fitted" = "#FF8C00")) +
  labs(title = "<span style='color:#4682B4'><b>Actual</b></span> & 
       <span style='color:#FF8C00'><b>Fitted</b></span> Electricity Demand", 
       x = NULL, y = NULL, color = "Variable") +
  geom_vline(xintercept = as.numeric(first_test_date), linetype = "dotted", color = "red", size = 1) +  # Add red dotted line
  theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10))
actual_predicted_plot

#------------Filtered Residual Plots------------
residual_plot <- ggplot(data_df , aes(x = Time)) +
  geom_line(aes(y = .resid, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  add_resid_points_func(list_of_dfs, color_string, # Generate geom_point layers
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 1)) +
  scale_color_manual(values = c("Actual" = "#00BFC4")) +
  labs(title = "Errors", x = NULL, y = NULL, color = "Variable") + 
  geom_vline(xintercept = as.numeric(first_test_date), linetype = "dotted", 
             color = "red", size = 1) +  # Add red dotted line
  theme_bw() + theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10)) 
residual_plot

two_plots <- grid.arrange(actual_predicted_plot, residual_plot, ncol = 1)
# Export the plot
ggsave("Electricity5.png", 
       plot = two_plots, width = 6, height = 4.5, dpi = 800) 
```

Calculate Precision and Related Metrics 
```{r}
#Calculate Precision for Train Set
train_filtered <- heart_merged_with_leaf %>%
  filter(Time >= as.POSIXct(train_start, tz = "UTC") & 
         Time <  as.POSIXct(train_end,   tz = "UTC"))

flagged_times <- bind_rows(list_of_dfs) %>%
  filter(Time >= as.POSIXct(train_start, tz = "UTC") & 
         Time <  as.POSIXct(train_end,   tz = "UTC")) %>%
  pull(Time)

train_metrics <- train_filtered %>%
  mutate(
    is_actual_extreme = (.resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
    is_flagged = Time %in% flagged_times,

    true_positive  = is_actual_extreme & is_flagged,
    false_positive = !is_actual_extreme & is_flagged
  )

tp <- sum(train_metrics$true_positive)
fp <- sum(train_metrics$false_positive)

precision <- tp / (tp + fp)

cat("\n--- Train Set Performance ---\n")
cat("Total flagged points:", tp + fp, "\n")
cat("Correctly identified extremes (TP):", tp, "\n")
cat("Incorrectly flagged (FP):", fp, "\n")
cat("Precision:", round(precision * 100, 2), "%\n")

# Calculate Precision for Test Set

test_filtered <- heart_merged_with_leaf %>%
  filter(Time >= as.POSIXct(test_start, tz = "UTC") & 
         Time <  as.POSIXct(test_end,   tz = "UTC"))

flagged_times_test <- bind_rows(list_of_dfs) %>%
  filter(Time >= as.POSIXct(test_start, tz = "UTC") & 
         Time <  as.POSIXct(test_end,   tz = "UTC")) %>%
  pull(Time)

test_metrics <- test_filtered %>%
  mutate(
    is_actual_extreme = (.resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
    is_flagged = Time %in% flagged_times_test,

    true_positive  = is_actual_extreme & is_flagged,
    false_positive = !is_actual_extreme & is_flagged
  )

tp_test <- sum(test_metrics$true_positive)
fp_test <- sum(test_metrics$false_positive)

precision_test <- tp_test / (tp_test + fp_test)

cat("\n--- Test Set Performance ---\n")
cat("Total flagged points:", tp_test + fp_test, "\n")
cat("Correctly identified extremes (TP):", tp_test, "\n")
cat("Incorrectly flagged (FP):", fp_test, "\n")
cat("Precision:", round(precision_test * 100, 2), "%\n")
```
```{r}

```




