---
title: "Eval Tree Analysis Code"
author: "Bobea"
date: "2024-02-03"
output: html_document
---
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      TreeAlert Heart Rate Example                         
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
Libraries
```{r, include=FALSE, warning=FALSE, message=FALSE}
library(readr); library(tidyr); library(feasts) # Data
library(tsibble); library(fable); library(fabletools) # Forecasting
library(dplyr); library(lubridate); library(gains)# Wrangling
library(ggplot2); library(ggtext); library(grid); library(gridExtra) # Plotting
library(slider); library(scales)
library(rpart); library(rpart.plot); # library(caret)

# ============================================================
# Load data
# ============================================================
heart_data_1 <- readRDS("~/Research/Eval Tree/Heart_Scripts/heart_combined.rds")
heart_data_2 <- readRDS("~/Research/Eval Tree/Heart_Scripts/heart_new_data.rds")
```
# Select analysis  read codeif you decide you prefer percentile analysis
```{r}
# ============================================================
# Merge metrics from heart_data_2 into heart_data_1
# ============================================================

heart_merged_clean <- heart_data_1 %>%
  left_join(
    heart_data_2 %>%
      select(
        Time,
        crps,
        winkler_score,
        lower,
        upper
      ),
    by = "Time"
  )

# ============================================================
# ANALYSIS TARGET SWITCH
# ============================================================

# Choose analysis target:
# 1 = original .resid (default)
# 2 = winkler_score
# 3 = crps
ANALYSIS_CHOICE <- 2

if (ANALYSIS_CHOICE == 1) {

  message("Analysis target: original .resid (no changes)")

} else if (ANALYSIS_CHOICE == 2) {

  message("Analysis target: winkler_score → .resid")

  heart_merged_clean <- heart_merged_clean %>%
    select(-.resid) %>%
    rename(.resid = winkler_score)

} else if (ANALYSIS_CHOICE == 3) {

  message("Analysis target: crps → .resid")

  heart_merged_clean <- heart_merged_clean %>%
    select(-.resid) %>%
    rename(.resid = crps)

} else {
  stop("Invalid ANALYSIS_CHOICE. Use 1 (.resid), 2 (winkler), or 3 (crps).")
}
#.resid will be replaces with your choice but keep the same column name! 

# ============================================================
# Train / Test split (85 / 15)
# ============================================================
split_index <- floor(0.85 * nrow(heart_merged_clean))

train_set <- heart_merged_clean[1:split_index, ]
test_set  <- heart_merged_clean[(split_index + 1):nrow(heart_merged_clean), ]
```
Tree Alert Functions:
```{r}
# To generate geom_point layers for 'All Filters Actual Vs Predicted'
add_fitted_points_func <- function(data_list, color_list, shape_values) {
  layers <- lapply(seq_along(data_list), function(i) {
    geom_point(data = data_list[[i]], 
               aes(x = Time, y = Value, shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
               size = 2.5, stroke = 1, color = color_list[i])
  })
  return(layers)
}

# To generate geom_point layers for 'All Filters Residuals'
add_resid_points_func <- function(data_list, color_list, shape_values) {
  layers <- lapply(seq_along(data_list), function(i) {
    geom_point(data = data_list[[i]], 
               aes(x = Time, y = .resid, shape = .resid >= top_5_percent_value | 
                     .resid <= bottom_5_percent_value),
               size = 2.5, stroke = 1, color = color_list[i])
  })
  return(layers)
}

# Function to create the plot for a specific month
create_monthly_plot <- function(data, start_date, end_date) {
  plot <- ggplot(data, aes(x = Time)) +
    geom_line(aes(y = Value, color = "Actual"), show.legend = FALSE) +
    geom_line(aes(y = model_1, color = "Fitted"), alpha = 0.6, show.legend = FALSE) +
    add_fitted_points_func(list_of_dfs, color_string, # make geom_point layers
                                      c("TRUE" = 1, "FALSE" = 16)) + 
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
    scale_color_manual(values = c("Actual" = "#4682B4", "Fitted" = "#FF8C00")) +
    scale_x_datetime(limits = as.POSIXct(c(start_date, end_date)), 
                     date_labels = "%m-%d",
                     #date_breaks = "3 days",
                     expand = c(0.02, 0.02)) +
    labs(title = paste(""), x = NULL, y = NULL, color = "Variable") +
    theme_bw() +
    theme(legend.position = "none",
          plot.title = element_markdown(size = 10),
          plot.title.position = "panel")
  
  return(plot)
}

# ----- Lift Functions -----

compute_lift_numeric <- function(df, actual_col, pred_col, n_groups = 50) {
    
  actual <- abs(df[[actual_col]])
  score  <- abs(df[[pred_col]])
  
  # Sort by predicted risk
  ord        <- order(score, decreasing = TRUE)
  actual_ord <- actual[ord]
  n          <- length(actual_ord)
  
  # Ventile grouping
  group <- ceiling(seq_len(n) / (n / n_groups))
  group[group > n_groups] <- n_groups
  
  df_lift <- data.frame(group = group, error = actual_ord) %>%
    dplyr::group_by(group) %>%
    dplyr::summarise(meanError = mean(error, na.rm = TRUE), .groups = "drop")
  
  # Normalize
  df_lift$meanError <- df_lift$meanError / mean(actual_ord, na.rm = TRUE)
  
  # Return first ventile lift
  df_lift$meanError[df_lift$group == 1]
}

make_lift_charts <- function(df, actual_col, pred_col, title_prefix = "Model") {
  
  actual <- abs(df[[actual_col]])
  score  <- abs(df[[pred_col]])
  
  # Sort by predicted risk
  ord <- order(score, decreasing = TRUE)
  actual_ord <- actual[ord]
  n <- length(actual_ord)
  
  # Cumulative gains
  df_cum <- data.frame(
    ncases  = c(0, seq_len(n)),
    cumError = c(0, cumsum(actual_ord))
  )
  
  # Ventiles
  n_groups <- 50
  group <- ceiling(seq_len(n) / (n / n_groups))
  group[group > n_groups] <- n_groups
  
  df_lift <- data.frame(
    group = group,
    error = actual_ord
  ) %>%
    dplyr::group_by(group) %>%
    dplyr::summarise(meanError = mean(error), .groups = "drop")
  
  df_lift$percentile <- df_lift$group * 5
  df_lift$meanError  <- df_lift$meanError / mean(actual_ord)
  
  g1 <- ggplot(df_cum, aes(ncases, cumError)) +
    geom_line(color = "#4682B4") +
    geom_abline(slope = sum(actual_ord)/n, intercept = 0,
                linetype = "dashed", color = "gray") +
    labs(title = paste(title_prefix, "– Cumulative Gains"),
         x = "# Cases", y = "Cumulative Error") +
    theme_minimal()
  
  g2 <- ggplot(df_lift, aes(percentile, meanError)) +
    geom_col(fill = "#FF8C00") +
    labs(title = paste(title_prefix, "– Ventile Lift"),
         x = "Percentile", y = "Ventile Mean / Global Mean") +
    theme_minimal()
  
  gridExtra::grid.arrange(g1, g2, ncol = 2)
  
  df_lift$meanError[df_lift$percentile == 5]
}
```
Train train/test split
```{r, warning=FALSE}
#------------Calculate Theta - Top / bottom 5% of data ------------
top_5_percent <- quantile(train_set$.resid, probs = 0.95, na.rm = TRUE)
bottom_5_percent <- quantile(train_set$.resid, probs = 0.05, na.rm = TRUE)
top_5_percent_value <- top_5_percent[[1]] 
bottom_5_percent_value <- bottom_5_percent[[1]]
top_5_percent_value ; bottom_5_percent_value 
```
# Regression Tree
```{r}
#------------Regression tree on .resid + temporal variables------------ 
tree_model <- rpart(.resid ~ Minute + Hour + Weekday + Intensity + Steps_10, 
                    data = train_set, method = "anova", 
                    cp = 0.00000, minbucket = 10)
tree_rules <- rpart.rules(tree_model, cover = FALSE, digits = 4, nn = FALSE)
#Attached the predicted tree nodes to DF and Maps rules/leafs to Observations

leaf_nodes <- predict(tree_model, newdata = heart_merged_clean, type = "vector") 
heart_merged_with_leaf <- dplyr::bind_cols(heart_merged_clean, Leaf_Node = leaf_nodes)
```
Lift 
```{r}
compute_lift <- function(data, model, pred_col_name, label, plot = TRUE) {
  
  # Predict leaf values
  leaf_nodes <- predict(model, newdata = data, type = "vector")
  data_with_pred <- dplyr::bind_cols(data, !!pred_col_name := leaf_nodes)
  
  lift_value <- if (plot) {
    make_lift_charts(
      df         = data_with_pred,
      actual_col = ".resid",
      pred_col   = pred_col_name,
      title_prefix = label
    )
  } else {
    compute_lift_numeric(
      df         = data_with_pred,
      actual_col = ".resid",
      pred_col   = pred_col_name
    )
  }
  
  cat(label, "Lift (First Quantile):", lift_value, "\n")
  
  return(list(
    lift = lift_value,
    merged_data = data_with_pred
  ))
}

train_results <- compute_lift(train_set, tree_model, "Leaf_Node", "Baseline (Train)")
test_results  <- compute_lift(test_set, tree_model, "Leaf_Node", "Baseline (Test)")
full_results  <- compute_lift(heart_merged_clean, tree_model, "Leaf_Node", "Baseline (Full)")
```

## Grid Search: Joint Optimization of Tree Parameters + Predictor Variables  ##
```{r}
## ============================================================================
## GRID SEARCH: Optimize Tree Parameters + Predictor Variables
## Using the SAME Lift Metric as make_lift_charts (Ventile Mean / Global Mean)
## ============================================================================

# 1. Candidate predictors
candidate_vars <- c(
  "Hour", "Minute", "Weekday", #"WeekofMonth", "DayofMonth",
  "Intensity", "Steps_10"
)

# 2. ALL non-empty combinations of predictors
var_combinations <- unlist(
  lapply(1:length(candidate_vars), function(k)
    combn(candidate_vars, k, FUN = paste, collapse = " + "))
)

# 3. Parameter grid
param_grid <- expand.grid(
  cp        = c(0, 0.0001, 0.001),
  minbucket = c(5, 10, 20),
  maxdepth  = c(5, 10, 15),
  vars      = var_combinations,
  stringsAsFactors = FALSE
)

# ----------------------------------------------------------------------------
# MODEL EVALUATION FUNCTION
# Uses compute_lift_numeric() exactly as your main lift code
# ----------------------------------------------------------------------------
evaluate_params <- function(params_row) {

  cp        <- params_row$cp
  minbucket <- params_row$minbucket
  maxdepth  <- params_row$maxdepth
  vars      <- params_row$vars

  # Build formula dynamically
  fmla <- as.formula(paste(".resid ~", vars))

  # Train regression tree
  model <- rpart(
    fmla,
    data = train_set,
    method = "anova",
    control = rpart.control(
      cp        = cp,
      minbucket = minbucket,
      maxdepth  = maxdepth
    )
  )

  # Predict on TRAIN SET
  preds <- predict(model, newdata = train_set)

  # Build temporary dataset to pass into compute_lift_numeric (df version)
  tmp_df <- dplyr::bind_cols(train_set, Pred_Temp = preds)

  # Compute top-ventile lift using SAME function as main analysis
  lift_value <- compute_lift_numeric(
    df       = tmp_df,
    actual_col = ".resid",
    pred_col   = "Pred_Temp",
    n_groups = 20
  )

  # Return row of results
  data.frame(
    cp                 = cp,
    minbucket          = minbucket,
    maxdepth           = maxdepth,
    vars               = vars,
    first_ventile_lift = lift_value
  )
}

# RUN GRID SEARCH
results <- do.call(
  rbind,
  lapply(seq_len(nrow(param_grid)), function(i) {
    evaluate_params(param_grid[i, , drop = FALSE])
  })
)

# SELECT BEST MODEL
best_params <- results %>%
  dplyr::filter(!is.na(first_ventile_lift)) %>%
  dplyr::arrange(dplyr::desc(first_ventile_lift)) %>%
  dplyr::slice(1)

cat("\n=============================\n")
cat("  BEST TREE PARAMETERS FOUND\n")
cat("=============================\n")
print(best_params)
```
Train Final Model using Best Parameters
```{r}
###############################################################################
## Train Final Model using Best Parameters + Best Variable Combination
###############################################################################

# Build formula dynamically from best_params$vars
final_formula <- as.formula(paste(".resid ~", best_params$vars))

cat("\n=============================\n")
cat("  FINAL MODEL FORMULA USED\n")
cat("=============================\n")
print(final_formula)

# Train final rpart model
best_tree_model <- rpart(
  final_formula,
  data = train_set,
  method = "anova",
  control = rpart.control(
    cp        = best_params$cp,
    minbucket = best_params$minbucket,
    maxdepth  = best_params$maxdepth
  )
)

# Display tree rules
cat("\n=============================\n")
cat("        FINAL TREE RULES\n")
cat("=============================\n")
tree_rules <- rpart.rules(best_tree_model, cover = FALSE, digits = 4, nn = FALSE)
#print(tree_rules)

# Attach predicted leaf node averages to full dataset
leaf_nodes <- predict(best_tree_model, newdata = heart_merged_clean, type = "vector")
heart_merged_with_leaf <- dplyr::bind_cols(heart_merged_clean, Leaf_Node = leaf_nodes)

# Preview result
cat("\n=============================\n")
cat(" ATTACHED LEAF NODE PREDICTIONS\n")
cat("=============================\n")
print(head(heart_merged_with_leaf))
```
# LIFT CHARTS FOR BEST GRID SEARCH MODEL (Train / Test / Full)
```{r}
# Reuse the SAME compute_lift function
compute_lift <- function(data, model, pred_col_name, label) {
  
  # Predict leaf node values
  leaf_nodes <- predict(model, newdata = data, type = "vector")
  
  # Create a merged dataset (does NOT modify original)
  data_with_pred <- dplyr::bind_cols(data, !!pred_col_name := leaf_nodes)
  
  # Compute lift
  lift_value <- make_lift_charts(
    df         = data_with_pred,
    actual_col = ".resid",
    pred_col   = pred_col_name,
    title_prefix = label
  )
  
  # Print and return results
  cat(label, "Lift (First Ventile):", lift_value, "\n")
  return(list(lift = lift_value, merged_data = data_with_pred))
}

# ---- Compute lift for BEST MODEL ----
best_train   <- compute_lift(train_set,            best_tree_model, "Pred_Best", "Best Model (Train)")
best_test    <- compute_lift(test_set,             best_tree_model, "Pred_Best", "Best Model (Test)")
best_full    <- compute_lift(heart_merged_clean,   best_tree_model, "Pred_Best", "Best Model (Full)")
```
Map rules to observations and  pull top 10 rules
```{r}
# How many rows of tree rules would you like to use?
start_row <- 1 # starting row of rule from top_resid_rules
end_row <- 3   # ending row (should be <= 20). if last chart does not # # Ensure it's a standard data frame

# ------------Pull top 10 rules------------
tree_rules$.resid <- as.numeric(tree_rules$.resid)
tree_rules <- tree_rules[order(abs(tree_rules$.resid),
                                      decreasing = TRUE), ]

top_10_resid_list <- as.list(tree_rules$.resid) # not top 10 now
last_column_index <- ncol(top_10_resid_list)
color_string <- c("#9400D3", "#008000", "#0000FF", "#FF8C00", "#000066",
                  "#FF00FF", "#AB0000", "#FF69B4", "#7B68EE", "#Ff0000", 
                  "#7B68EE", "#20B2AA", "#703366", "#32CD32", "#CC6633", 
                  "#C77CFF", "#8A2BE2", "#00FF7F", "#00BFFF", "#FFA500", 
                  "#66CDAA", "#BA55D3", "#00CED1", "#4682B4", "#00FFFF")
```

# THE PROGRAM -- appllies rules as filter and plots the results
```{r, warning=FALSE, message=FALSE}
#------------Filter the rows for each rule for plotting------------
list_of_dfs <- list() # Create an empty DataFrame to store the results
for (i in start_row:end_row) {

  # Pull the next iterative resid to filter by form the top 10 list
  current_rule <- as.numeric(top_10_resid_list[[i]])
  # round both to common decimal as heart_merged_with_leaf has 1 more decimal place
  precision <- 3  
  # Filter heart_merged_with_leaf
  filtered_data <- heart_merged_with_leaf %>%
    filter(round(Leaf_Node, precision) == round(current_rule, precision))
  #------------displays the rule text before each plot------------
  first_row <- tree_rules[i, ]
  non_blank_cells <- first_row[first_row != ""]
  constructed_sentence <- paste(names(non_blank_cells), non_blank_cells, collapse = " ")
    rule_text <- textGrob(constructed_sentence,
                           gp = gpar(col = "black", fontsize = 9))
    print(constructed_sentence)
  #------------Plot of filtered rules on fitted time scale------------
  post_rules_scatter_plot <- ggplot(filtered_data, 
                                    aes(x = Time, y = .resid)) + 
    geom_point(aes(color = .resid > 10)) +
    geom_text(aes(label = format(Time, "%H:%M")), 
              hjust = -0.1, vjust = 0.5, size = 2) +  # Adding time labels
    labs(title = "Scatter Plot of Residuals", x = "Time", y = "Residual") +
      scale_color_manual(values = c("FALSE" = "#FF0000", "TRUE" = "#9400D3")) +
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 2)) 
  #------------Create Residual Plot------------
  #Initiate arguments
  #shape_number <- ifelse(row <= 25, row, 20)
  random_color <- rgb(runif(1), runif(1), runif(1), maxColorValue = 255)
  list_of_dfs[[i]] <- filtered_data # used to save df's for lower plot.
  # Plot of filtered rules on full series
  post_rules_line_plot <- ggplot(heart_merged_clean, 
                                 aes(x = Time, y = .resid)) +
    geom_line(color = "#00BFC4") +
    geom_point(data = filtered_data, 
               aes(x = Time, y = .resid,
                   color = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
                   shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
                   size = 2.5, stroke = 0.8) +
    scale_color_manual(values = c("FALSE" = color_string[i], 
                                  "TRUE" = color_string[i])) +  
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
    labs(title = paste("Residual Plot with Tree Filtered Extreme and Non-extreme Values",
                       sep=""), x = NULL, y = NULL) +
    geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", linetype = "dashed") +
    theme_bw() +
    theme(legend.position = "none") + 
    theme(plot.title = element_markdown())
    #------------Plot of Actual Vs Predicted------------
  actual_predicted_plot <- ggplot(heart_merged_clean, aes(x = Time)) +
    geom_line(aes(y = Value, color = "Actual"), show.legend = FALSE) +
    geom_point(data = filtered_data, 
               aes(x = Time, y = Value, 
      shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
      color = .resid >= top_5_percent_value | .resid <=bottom_5_percent_value),
                   size = 2.5, stroke = 1, show.legend = FALSE) +  
    scale_color_manual(values = c("TRUE" = color_string[i], 
                                  "FALSE" = color_string[i], 
                                  "Actual" = "#4682B4"),
                       guide = FALSE) + 
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +  
    labs(title = "<span style='color:#4682B4'><b>Actual</b>
         </span> Heart Rates + Tree Filtered Points",
         x = NULL, y = NULL, color = "Variable") +
    theme(legend.position = "none") +
    geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", linetype = "dashed") +
    theme_bw() +
    theme(plot.title = element_markdown())

   # Output plots
    grid.arrange(rule_text, post_rules_scatter_plot, ncol = 1)
    grid.arrange(post_rules_line_plot, actual_predicted_plot,
                 ncol = 1, heights = c(10,10))
}
```

```{r, warning=FALSE, message=FALSE}
#fills gaps with NA's so that the line does not appear on blanks. 
heart_merged_clean_filled <- heart_merged_clean %>%
  complete(Time = seq(min(Time), max(Time), by = "10 min"))
# -----------
# Filter points based on your criteria
highlight_points <- subset(heart_merged_clean_filled,
             (Hour %in% c("5", "6") & Weekday %in% 
                c("1", "2", "3", "4", "5")) |  # Weekday early AM
             Weekday %in% c("7") |  # Weekends
             WeekofMonth %in% c("1", "4", "5")) # 1st/4th/5th weeks of the month

# Plot of filtered rules on full series
post_rules_line_plot <- ggplot(heart_merged_clean_filled, aes(x = Time, y = .resid)) +
  geom_line(color = "#00BFC4") +
  geom_point(data = highlight_points, 
             aes(x = Time, y = .resid, color = .resid > 10), size = 0.5) +  
  labs(title = "Residual Plot: <span style='color:#9900D0'><b>Extreme</b></span> vs <span 
       style='color:#F8666D'><b>Non-extreme</b></span> values",
       x = NULL, y = NULL, color = "Variable") +
  scale_color_manual(values = c("FALSE" = "#F8766D", "TRUE" = "#C77CFF")) +
  scale_x_datetime(date_breaks = "4 day", date_labels = "%m-%d", 
                   expand = c(0.02, 0.02)) +
  theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown())

#------------All Filters Actual Vs Predicted------------
# Create the plot
actual_predicted_plot <- ggplot(heart_merged_clean_filled, aes(x = Time)) +
  geom_line(aes(y = Value, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  geom_line(aes(y = .fitted, color = "Fitted"), 
            alpha = 0.6, show.legend = FALSE) +
  add_fitted_points_func(list_of_dfs, color_string, 
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 1)) +
  scale_color_manual(values = c("Actual" = "#4682B4", "Fitted" = "#FF8C00")) +
  labs(title = "<span style='color:#4682B4'><b>Actual</b></span> & 
       <span style='color:#FF8C00'><b>Fitted</b></span> Heart Rates", 
       x = NULL, y = NULL, color = "Variable") +
  geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", 
             linetype = "dotted", size = 1) +
  theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10))

actual_predicted_plot

# Create the plot
residual_plot <- ggplot(heart_merged_clean_filled, aes(x = Time)) +
  geom_line(aes(y = .resid, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  add_resid_points_func(list_of_dfs, color_string, # Generate geom_point layers
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 1)) +
  scale_color_manual(values = c("Actual" = "#00BFC4")) +
  labs(title = "Errors", x = NULL, y = NULL, color = "Variable") +
  geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", 
             linetype = "dotted", size = 1) + theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10)) 
residual_plot

two_plots <- grid.arrange(actual_predicted_plot, residual_plot, ncol = 1)
# Export the plot
#ggsave("HeartRate.png", 
#       plot = two_plots, width = 6, height = 4, dpi = 800) 
ggsave("hr_viz.png", 
       plot = two_plots, width = 6, height = 4.5, dpi = 800)
```
Misc
```{r}
percent_above_10 <- mean(highlight_points$.resid > 10) * 100
average_resid <- mean(highlight_points$.resid)
cat("Percentage of values above 10:", percent_above_10, "%\n")
cat("Average value of .resid column:", average_resid, "\n")
```
Grid search
```{r}
# # Define grid of parameters
# cp_vals <- c(0, 0.0001, 0.001)
# minbucket_vals <- c(5, 10, 20)
# maxdepth_vals <- c(5, 10, 15)
# 
# param_grid <- expand.grid(cp = cp_vals,
#                           minbucket = minbucket_vals,
#                           maxdepth = maxdepth_vals)
# 
# # Safe evaluation function
# eval_tree_lift <- function(cp, minbucket, maxdepth, train_set) {
#   model <- rpart(.resid ~ Minute + Hour + Weekday + Intensity + Steps_10,
#                  data = train_set,
#                  method = "anova",
#                  control = rpart.control(cp = cp,
#                                          minbucket = minbucket,
#                                          maxdepth = maxdepth))
#   
#   # Predict fitted values
#   train_set$.fitted <- predict(model, newdata = train_set)
#   actual_error <- abs(train_set$.resid)
#   predicted_error <- train_set$.fitted
#   
#   # Compute gains
#   gain <- tryCatch({
#     gains(actual_error, predicted_error, groups = 20) #changes decile to 20
#   }, error = function(e) {
#     return(NULL)
#   })
#   
#   # Extract first decile lift safely
#   if (!is.null(gain) && 10 %in% gain$depth) {
#     lift <- gain$mean.resp[gain$depth == 10] / mean(actual_error)
#     return(lift)
#   } else {
#     return(NA_real_)
#   }
# }
# 
# # Run grid search
# results <- param_grid %>%
#   rowwise() %>%
#   mutate(first_decile_lift = list(eval_tree_lift(cp, minbucket, maxdepth, heart_merged_with_leaf))) %>%
#   unnest(cols = c(first_decile_lift)) %>%
#   ungroup()
# 
# # Find best parameters
# best_params <- results %>% arrange(desc(first_decile_lift)) %>% slice(1)
```
