---
title: "Eval Tree Analysis Code"
author: "Bobea"
date: "2024-02-03"
output: html_document
---
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      TreeAlert Heart Rate Example                         
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
Libraries
```{r, include=FALSE, warning=FALSE, message=FALSE}
library(readr); library(tidyr); library(feasts) # Data
library(tsibble); library(fable); library(fabletools) # Forecasting
library(dplyr); library(lubridate) # Wrangling
library(ggplot2); library(ggtext); library(grid); library(gridExtra) # Plotting
library(slider); library(scales); library(rpart); library(rpart.plot); 
source(file.path(if (dir.exists("TA_functions")) "." else "..", "TA_functions", "TA_functions.R"))
source(file.path(if (dir.exists("TA_functions")) "." else "..", "TA_functions", "Rephrase_Rule_Func.R"))
# library(caret)
```

Configuration
```{r}
# Choose analysis target:
# 1 = point-forecast residuals optimized with RMSE lift
# 2 = winkler_score
# 3 = crps
ANALYSIS_CHOICE <- 1

n_groups <- 20
n_rules <- 3
minute_increment <- 10

baseline_formula <- .resid ~ Minute + Hour + Weekday + Intensity + Steps_10
fixed_formula <- .resid ~ Hour + Minute + Weekday + Intensity + Steps_10
final_formula <- fixed_formula

param_grid <- expand.grid(
  cp        = c(0, 0.0001, 0.001, 0.005, 0.01),
  minbucket = seq(4, 8, by = 1),
  maxdepth  = seq(1, 6, by = 1)
)
```

Load Data
```{r, include=FALSE, warning=FALSE, message=FALSE}
heart_dir <- if (dir.exists("Heart_Rate")) "Heart_Rate" else "."
heart_image_dir <- file.path(heart_dir, "images")
heart_sensitivity_image_dir <- file.path(heart_image_dir, "sensitivity")
heart_output_prefix <- dplyr::case_when(
  ANALYSIS_CHOICE == 1 ~ "hr_rmse",
  ANALYSIS_CHOICE == 2 ~ "hr_winkler_score",
  ANALYSIS_CHOICE == 3 ~ "hr_crps",
  TRUE ~ "hr_analysis"
)
heart_image_file <- function(filename) {
  paste0(heart_output_prefix, "_", filename)
}
dir.create(heart_image_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(heart_sensitivity_image_dir, showWarnings = FALSE, recursive = TRUE)
heart_data_1 <- readRDS(file.path(heart_dir, "HR_data", "heart_combined.rds"))
heart_data_2 <- readRDS(file.path(heart_dir, "HR_data", "heart_new_data.rds"))
```

Prepare Data
```{r}
# Merge metrics from heart_data_2 into heart_data_1

analysis_data <- heart_data_1 %>%
  left_join(
    heart_data_2 %>%
      select(
        Time,
        crps,
        winkler_score,
        lower,
        upper
      ),
    by = "Time"
  )

if (ANALYSIS_CHOICE == 1) {
  message("Analysis target: original point-forecast residuals with RMSE lift")
  analysis_target_label <- "Forecast Error"
  lift_metric <- "rmse"
} else if (ANALYSIS_CHOICE == 2) {
  message("Analysis target: winkler_score -> .resid")
  analysis_target_label <- "Winkler Score"
  lift_metric <- "mean_abs"
  analysis_data <- analysis_data %>%
    select(-.resid) %>%
    rename(.resid = winkler_score)
} else if (ANALYSIS_CHOICE == 3) {
  message("Analysis target: crps -> .resid")
  analysis_target_label <- "CRPS"
  lift_metric <- "mean_abs"
  analysis_data <- analysis_data %>%
    select(-.resid) %>%
    rename(.resid = crps)
} else {
  stop("Invalid ANALYSIS_CHOICE. Use 1 (RMSE), 2 (winkler), or 3 (crps).")
}
#.resid will be replaced with your choice but keep the same column name!

# ============================================================
# Train / Test split (85 / 15)
split_index <- floor(0.85 * nrow(analysis_data))

train_set <- analysis_data[1:split_index, ]
test_set  <- analysis_data[(split_index + 1):nrow(analysis_data), ]
```
Train/Test Split
```{r, warning=FALSE}
#------------Calculate Theta - Top / bottom 5% of data ------------
top_5_percent <- quantile(train_set$.resid, probs = 0.95, na.rm = TRUE)
bottom_5_percent <- quantile(train_set$.resid, probs = 0.05, na.rm = TRUE)
top_5_percent_value <- top_5_percent[[1]] 
bottom_5_percent_value <- bottom_5_percent[[1]]
top_5_percent_value ; bottom_5_percent_value 
```

Baseline Tree
```{r}
#------------Regression tree on .resid + temporal variables------------ 
tree_model <- rpart(baseline_formula, 
                    data = train_set, method = "anova", 
                    cp = 0.00000, minbucket = 10)
tree_rules <- rpart.rules(tree_model, cover = FALSE, digits = 4, nn = FALSE)
#Attached the predicted tree nodes to DF and Maps rules/leafs to Observations

leaf_nodes <- predict(tree_model, newdata = analysis_data, type = "vector") 
analysis_data_with_leaf <- dplyr::bind_cols(analysis_data, Leaf_Node = leaf_nodes)
```
Lift Charts
```{r}
train_results <- tree_alert_compute_lift(
  train_set, tree_model, "Leaf_Node", "Baseline (Train)",
  plot = TRUE, n_groups = n_groups,
  lift_label = "First Quantile",
  lift_metric = lift_metric,
  axis_title_size = 20, axis_text_x_size = 18, axis_text_y_size = 20
)
test_results <- tree_alert_compute_lift(
  test_set, tree_model, "Leaf_Node", "Baseline (Test)",
  plot = TRUE, n_groups = n_groups,
  lift_label = "First Quantile",
  lift_metric = lift_metric,
  axis_title_size = 20, axis_text_x_size = 18, axis_text_y_size = 20
)
```

Grid Search
```{r}
## ============================================================================
## GRID SEARCH: Optimize Tree Parameters ONLY
## Fixed Predictors:
## Hour + Minute + Weekday + Intensity + Steps_10
## Using SAME Lift Metric
## ============================================================================

# RUN GRID SEARCH
grid_search <- tree_alert_grid_search(
  train_set = train_set,
  formula = fixed_formula,
  param_grid = param_grid,
  n_groups = n_groups,
  pred_col_name = "Pred_Temp",
  lift_metric = lift_metric
)

# SELECT BEST MODEL
results <- grid_search$results
best_params <- grid_search$best_params

cat("\n=============================\n")
cat("  BEST TREE PARAMETERS FOUND\n")
cat("=============================\n")
print(best_params)
```

Final Model
```{r}
cat("  FINAL MODEL FORMULA USED\n")
print(final_formula)

# Train final rpart model
best_tree_model <- tree_alert_train_best_model(
  train_set = train_set,
  formula = final_formula,
  best_params = best_params
)

# Display tree rules
cat("\n=============================\n")
cat("        FINAL TREE RULES\n")
cat("=============================\n")
tree_rules <- rpart.rules(best_tree_model, cover = FALSE, digits = 4, nn = FALSE)
#print(tree_rules)

# Attach predicted leaf node averages to full dataset
leaf_nodes <- predict(best_tree_model, newdata = analysis_data, type = "vector")
analysis_data_with_leaf <- dplyr::bind_cols(analysis_data, Leaf_Node = leaf_nodes)

# Preview result
cat("\n=============================\n")
cat(" ATTACHED LEAF NODE PREDICTIONS\n")
cat("=============================\n")
print(head(analysis_data_with_leaf))
```
Final Lift Charts
```{r}
# ---- Compute lift for BEST MODEL ----
best_train <- tree_alert_compute_lift(
  train_set, best_tree_model, "Pred_Best", "Best Model (Train)",
  plot = TRUE, n_groups = n_groups,
  lift_label = "First Ventile",
  lift_metric = lift_metric,
  axis_title_size = 20, axis_text_x_size = 18, axis_text_y_size = 20
)
ggsave(file.path(heart_image_dir, heart_image_file("lift_train.png")), plot = last_plot(), width = 8, height = 6, dpi = 1000)
best_test <- tree_alert_compute_lift(
  test_set, best_tree_model, "Pred_Best", "Best Model (Test)",
  plot = TRUE, n_groups = n_groups,
  lift_label = "First Ventile",
  lift_metric = lift_metric,
  axis_title_size = 20, axis_text_x_size = 18, axis_text_y_size = 20
)
ggsave(file.path(heart_image_dir, heart_image_file("lift_test.png")), plot = last_plot(), width = 8, height = 6, dpi = 1000)
```

Top Rules
```{r}
# ------------Pull top sorted terminal-node rules------------
rule_selection <- tree_alert_pull_top_k_rules(
  tree_rules = tree_rules,
  n_rules = n_rules
)
tree_rules <- rule_selection$tree_rules
selected_rules <- rule_selection$selected_rules
selected_rule_values <- rule_selection$selected_rule_values
rule_text_output <- tree_alert_list_top_rules(
  selected_rules = selected_rules,
  minute_increment = minute_increment
)
```

Rule Plots
```{r, warning=FALSE, message=FALSE}
#------------Filter the rows for each rule for plotting------------
list_of_dfs <- list() # Create an empty DataFrame to store the results
rule_coverage <- data.frame()
for (i in seq_along(selected_rule_values)) {

  # Pull the next sorted rule value to filter observations for plotting
  current_rule <- as.numeric(selected_rule_values[[i]])
  # round both to common decimal as analysis_data_with_leaf has more precision
  precision <- 2
  # Filter analysis_data_with_leaf
  filtered_data <- analysis_data_with_leaf %>%
    as.data.frame() %>%
    filter(round(Leaf_Node, precision) == round(current_rule, precision))
  
  rule_coverage <- bind_rows(
    rule_coverage,
    data.frame(
      rule = i,
      train_points = sum(filtered_data$Time < min(test_set$Time, na.rm = TRUE)),
      test_points = sum(filtered_data$Time >= min(test_set$Time, na.rm = TRUE)),
      total_points = nrow(filtered_data)
    )
  )
  
  constructed_sentence <- rule_text_output$raw_rules[i]
  rule_text <- textGrob(constructed_sentence, gp = gpar(col = "black", fontsize = 9))
  
  #------------Plot of filtered rules on fitted time scale------------
  post_rules_scatter_plot <- ggplot(filtered_data, 
                                    aes(x = Time, y = .resid)) + 
    geom_point(aes(color = .resid > 10)) +
    geom_text(aes(label = format(Time, "%H:%M")), 
              hjust = -0.1, vjust = 0.5, size = 2) +  # Adding time labels
  labs(title = paste("Scatter Plot of", analysis_target_label), x = "Time", y = analysis_target_label) +
      scale_color_manual(values = c("FALSE" = "#FF0000", "TRUE" = "#9400D3"))
  #------------Create Residual Plot------------
  #Initiate arguments
  #shape_number <- ifelse(row <= 25, row, 20)
  random_color <- rgb(runif(1), runif(1), runif(1), maxColorValue = 255)
  list_of_dfs[[i]] <- filtered_data # used to save df's for lower plot.
  # Plot of filtered rules on full series
  post_rules_line_plot <- ggplot(analysis_data, 
                                 aes(x = Time, y = .resid)) +
    geom_line(color = "#00BFC4") +
    geom_point(data = filtered_data, 
               aes(x = Time, y = .resid,
                   color = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
                   shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
                   size = 2.5, stroke = 0.8) +
    scale_color_manual(values = c("FALSE" = color_string[i], 
                                  "TRUE" = color_string[i])) +  
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
    labs(title = paste(analysis_target_label, "Plot with Tree Filtered Extreme and Non-extreme Values",
                       sep=""), x = NULL, y = NULL) +
    geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", linetype = "dashed") +
    theme_bw() +
    theme(legend.position = "none") + 
    theme(plot.title = element_markdown())
    #------------Plot of Actual Vs Predicted------------
  actual_predicted_plot <- ggplot(analysis_data, aes(x = Time)) +
    geom_line(aes(y = Value, color = "Actual"), show.legend = FALSE) +
    geom_point(data = filtered_data, 
               aes(x = Time, y = Value, 
      shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
      color = .resid >= top_5_percent_value | .resid <=bottom_5_percent_value),
                   size = 2.5, stroke = 1, show.legend = FALSE) +  
    scale_color_manual(values = c("TRUE" = color_string[i], 
                                  "FALSE" = color_string[i], 
                                  "Actual" = "#4682B4"),
                       guide = FALSE) + 
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +  
    labs(title = "<span style='color:#4682B4'><b>Actual</b>
         </span> Heart Rates + Tree Filtered Points",
         x = NULL, y = NULL, color = "Variable") +
    theme(legend.position = "none") +
    geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", linetype = "dashed") +
    theme_bw() +
    theme(plot.title = element_markdown())

   # Output plots
    grid.newpage()
    grid.draw(arrangeGrob(rule_text, post_rules_scatter_plot, ncol = 1))
    grid.newpage()
    grid.draw(arrangeGrob(post_rules_line_plot, actual_predicted_plot,
                 ncol = 1, heights = c(10,10)))
}

print(rule_coverage)
```

All Rules Plots
```{r, warning=FALSE, message=FALSE}
#fills gaps with NA's so that the line does not appear on blanks. 
analysis_data_filled <- analysis_data %>%
  complete(Time = seq(min(Time), max(Time), by = "10 min"))
# -----------
# Filter points based on your criteria
highlight_points <- subset(analysis_data_filled,
             (Hour %in% c("5", "6") & Weekday %in% 
                c("1", "2", "3", "4", "5")) |  # Weekday early AM
             Weekday %in% c("7") |  # Weekends
             WeekofMonth %in% c("1", "4", "5")) # 1st/4th/5th weeks of the month

# Plot of filtered rules on full series
post_rules_line_plot <- ggplot(analysis_data_filled, aes(x = Time, y = .resid)) +
  geom_line(color = "#00BFC4") +
  geom_point(data = highlight_points, 
             aes(x = Time, y = .resid, color = .resid > 10), size = 0.5) +  
  labs(title = "Residual Plot: <span style='color:#9900D0'><b>Extreme</b></span> vs <span 
       style='color:#F8666D'><b>Non-extreme</b></span> values",
       x = NULL, y = NULL, color = "Variable") +
  scale_color_manual(values = c("FALSE" = "#F8766D", "TRUE" = "#C77CFF")) +
  scale_x_datetime(date_breaks = "4 day", date_labels = "%m-%d", 
                   expand = c(0.02, 0.02)) +
  theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown())

#------------All Filters Actual Vs Predicted------------
# Create the plot
actual_predicted_plot <- ggplot(analysis_data_filled, aes(x = Time)) +
  geom_line(aes(y = Value, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  geom_line(aes(y = .fitted, color = "Fitted"), 
            alpha = 0.6, show.legend = FALSE) +
  add_fitted_points_func(list_of_dfs, color_string, 
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 1)) +
  scale_color_manual(values = c("Actual" = "#4682B4", "Fitted" = "#FF8C00")) +
  labs(title = "<span style='color:#4682B4'><b>Actual</b></span> & 
       <span style='color:#FF8C00'><b>Fitted</b></span> Heart Rates", 
       x = NULL, y = NULL, color = "Variable") +
  geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", 
             linetype = "dotted", size = 1) +
  theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10))

actual_predicted_plot

# Create the plot
residual_plot <- ggplot(analysis_data_filled, aes(x = Time)) +
  geom_line(aes(y = .resid, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  add_resid_points_func(list_of_dfs, color_string, # Generate geom_point layers
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 1)) +
  scale_color_manual(values = c("Actual" = "#00BFC4")) +
  labs(title = analysis_target_label, x = NULL, y = NULL, color = "Variable") +   
  geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", 
             linetype = "dotted", size = 1) + theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10)) 
residual_plot

two_plots <- grid.arrange(actual_predicted_plot, residual_plot, ncol = 1)
# Export the plot
#ggsave("HeartRate.png", 
#       plot = two_plots, width = 6, height = 4, dpi = 800) 
ggsave(file.path(heart_image_dir, heart_image_file("viz.png")), 
       plot = two_plots, width = 6, height = 4.5, dpi = 800)
```
Metrics
```{r}
performance_summary <- tree_alert_performance_summary(
  list_of_dfs = list_of_dfs,
  train_set = train_set,
  test_set = test_set,
  train_lift = best_train$lift,
  test_lift = best_test$lift
)
```

Sensitivity Analysis
```{r, warning=FALSE, message=FALSE}
sensitivity_analysis <- tree_alert_sensitivity_analysis(
  train_set = train_set,
  test_set = test_set,
  formula = final_formula,
  param_grid = param_grid,
  n_groups = n_groups,
  lift_metric = lift_metric,
  reference_train_lift = best_train$lift,
  reference_test_lift = best_test$lift,
  train_image_path = file.path(heart_sensitivity_image_dir, heart_image_file("sensitivity_train_lift.png")),
  test_image_path = file.path(heart_sensitivity_image_dir, heart_image_file("sensitivity_test_lift.png"))
)

sensitivity_results <- sensitivity_analysis$results
sensitivity_train_plot <- sensitivity_analysis$train_plot
sensitivity_test_plot <- sensitivity_analysis$test_plot
sensitivity_train_plot
sensitivity_test_plot
```

Train/Test Split Sensitivity Analysis
```{r, warning=FALSE, message=FALSE}
split_sensitivity_analysis <- tree_alert_split_sensitivity_analysis(
  data = analysis_data,
  formula = final_formula,
  param_grid = param_grid,
  n_groups = n_groups,
  lift_metric = lift_metric,
  train_share_min = 70,
  train_share_max = 90,
  train_share_step = 1,
  reference_train_lift = best_train$lift,
  reference_test_lift = best_test$lift,
  train_image_path = file.path(heart_sensitivity_image_dir, heart_image_file("split_sensitivity_train_lift.png")),
  test_image_path = file.path(heart_sensitivity_image_dir, heart_image_file("split_sensitivity_test_lift.png"))
)

split_sensitivity_results <- split_sensitivity_analysis$results
split_sensitivity_train_plot <- split_sensitivity_analysis$train_plot
split_sensitivity_test_plot <- split_sensitivity_analysis$test_plot
split_sensitivity_train_plot
split_sensitivity_test_plot
```

Feature Granularity Sensitivity Analysis
```{r, warning=FALSE, message=FALSE}
feature_sensitivity_analysis <- tree_alert_feature_sensitivity_analysis(
  train_set = train_set,
  test_set = test_set,
  feature_formulas = list(
    "Hour" = .resid ~ Hour,
    "Minute + Hour" = .resid ~ Minute + Hour,
    "Hour + Weekday" = .resid ~ Hour + Weekday,
    "Minute + Hour + Weekday" = .resid ~ Minute + Hour + Weekday
  ),
  param_grid = param_grid,
  n_groups = n_groups,
  lift_metric = lift_metric,
  reference_train_lift = best_train$lift,
  reference_test_lift = best_test$lift,
  image_path = file.path(heart_sensitivity_image_dir, heart_image_file("feature_sensitivity_lift.png")),
  height = 4.2
)

feature_sensitivity_results <- feature_sensitivity_analysis$results
feature_sensitivity_plot <- feature_sensitivity_analysis$plot
feature_sensitivity_results
feature_sensitivity_plot
```

K-Rule Sensitivity Analysis
```{r, warning=FALSE, message=FALSE}
k_rule_sensitivity_analysis <- tree_alert_k_rule_sensitivity_analysis(
  train_set = train_set,
  test_set = test_set,
  tree_rules = tree_rules,
  model = best_tree_model,
  k_values = 1:10,
  n_groups = n_groups,
  lift_metric = lift_metric,
  image_path = file.path(heart_sensitivity_image_dir, heart_image_file("k_rule_sensitivity_lift.png"))
)

k_rule_sensitivity_results <- k_rule_sensitivity_analysis$results
k_rule_sensitivity_plot <- k_rule_sensitivity_analysis$plot
k_rule_sensitivity_results
k_rule_sensitivity_plot
```

Update Results Summary
```{r, warning=FALSE, message=FALSE}
heart_results_section_id <- dplyr::case_when(
  ANALYSIS_CHOICE == 1 ~ "heart_rate_rmse",
  ANALYSIS_CHOICE == 2 ~ "heart_rate_winkler_score",
  ANALYSIS_CHOICE == 3 ~ "heart_rate_crps",
  TRUE ~ "heart_rate_other"
)

heart_results_title <- dplyr::case_when(
  ANALYSIS_CHOICE == 1 ~ "Heart Rate: Point-Forecast Residuals (ANALYSIS_CHOICE = 1)",
  ANALYSIS_CHOICE == 2 ~ "Heart Rate: Winkler Score (ANALYSIS_CHOICE = 2)",
  ANALYSIS_CHOICE == 3 ~ "Heart Rate: CRPS (ANALYSIS_CHOICE = 3)",
  TRUE ~ paste("Heart Rate: Analysis Choice", ANALYSIS_CHOICE)
)

tree_alert_update_results_summary(
  output_path = file.path(if (dir.exists("TA_functions")) "." else "..", "Results", "TreeAlert_results_summary.md"),
  section_id = heart_results_section_id,
  title = heart_results_title,
  analysis_choice = ANALYSIS_CHOICE,
  analysis_target = analysis_target_label,
  formula = final_formula,
  n_groups = n_groups,
  n_rules = n_rules,
  minute_increment = minute_increment,
  train_set = train_set,
  test_set = test_set,
  best_params = best_params,
  train_lift = best_train$lift,
  test_lift = best_test$lift,
  performance_summary = performance_summary,
  rule_text_output = rule_text_output,
  rule_coverage = rule_coverage,
  sensitivity_results = sensitivity_results,
  split_sensitivity_results = split_sensitivity_results,
  feature_sensitivity_results = feature_sensitivity_results,
  k_rule_sensitivity_results = k_rule_sensitivity_results,
  image_outputs = data.frame(
    output = c(
      "train_lift", "test_lift", "visualization",
      "hyperparameter_train", "hyperparameter_test",
      "split_train", "split_test",
      "feature_granularity", "k_rule"
    ),
    path = c(
      file.path(heart_image_dir, heart_image_file("lift_train.png")),
      file.path(heart_image_dir, heart_image_file("lift_test.png")),
      file.path(heart_image_dir, heart_image_file("viz.png")),
      file.path(heart_sensitivity_image_dir, heart_image_file("sensitivity_train_lift.png")),
      file.path(heart_sensitivity_image_dir, heart_image_file("sensitivity_test_lift.png")),
      file.path(heart_sensitivity_image_dir, heart_image_file("split_sensitivity_train_lift.png")),
      file.path(heart_sensitivity_image_dir, heart_image_file("split_sensitivity_test_lift.png")),
      file.path(heart_sensitivity_image_dir, heart_image_file("feature_sensitivity_lift.png")),
      file.path(heart_sensitivity_image_dir, heart_image_file("k_rule_sensitivity_lift.png"))
    )
  )
)
```
