---
title: "Eval Tree Analysis Code"
author: "Bobea"
date: "2024-02-03"
output: html_document
---
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                      TreeAlert Heart Rate Example                         
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
Libraries
```{r, include=FALSE, warning=FALSE, message=FALSE}
library(readr); library(tidyr); library(feasts) # Data
library(tsibble); library(fable); library(fabletools) # Forecasting
library(dplyr); library(lubridate) # Wrangling
library(ggplot2); library(ggtext); library(grid); library(gridExtra) # Plotting
library(slider); library(scales); library(rpart); library(rpart.plot); 
source(file.path(if (dir.exists("TA_functions")) "." else "..", "TA_functions", "TA_functions.R"))
# library(caret)
# ============================================================
# Load data
heart_dir <- if (dir.exists("Heart_Rate")) "Heart_Rate" else "."
heart_image_dir <- file.path(heart_dir, "images")
dir.create(heart_image_dir, showWarnings = FALSE, recursive = TRUE)
heart_data_1 <- readRDS(file.path(heart_dir, "HR_data", "heart_combined.rds"))
heart_data_2 <- readRDS(file.path(heart_dir, "HR_data", "heart_new_data.rds"))
```
# DONT SKIP--> Select analysis.  Choose option 2 if you prefer percentile analysis
```{r}
# ============================================================
# ANALYSIS TARGET SWITCH
# ============================================================

# Choose analysis target:
# 1 = original .resid (default)
# 2 = winkler_score
# 3 = crps

ANALYSIS_CHOICE <- 2

#--------------------------
# Merge metrics from heart_data_2 into heart_data_1

heart_merged_clean <- heart_data_1 %>%
  left_join(
    heart_data_2 %>%
      select(
        Time,
        crps,
        winkler_score,
        lower,
        upper
      ),
    by = "Time"
  )

if (ANALYSIS_CHOICE == 1) {
  message("Analysis target: original .resid (no changes)")
} else if (ANALYSIS_CHOICE == 2) {
  message("Analysis target: winkler_score → .resid")
  heart_merged_clean <- heart_merged_clean %>%
    select(-.resid) %>%
    rename(.resid = winkler_score)
} else if (ANALYSIS_CHOICE == 3) {
  message("Analysis target: crps → .resid")
  heart_merged_clean <- heart_merged_clean %>%
    select(-.resid) %>%
    rename(.resid = crps)
} else {
  stop("Invalid ANALYSIS_CHOICE. Use 1 (.resid), 2 (winkler), or 3 (crps).")
}
#.resid will be replaces with your choice but keep the same column name! 

# ============================================================
# Train / Test split (85 / 15)
split_index <- floor(0.85 * nrow(heart_merged_clean))

train_set <- heart_merged_clean[1:split_index, ]
test_set  <- heart_merged_clean[(split_index + 1):nrow(heart_merged_clean), ]
```
Tree Alert Functions:
```{r}
# Function to create the plot for a specific month
create_monthly_plot <- function(data, start_date, end_date) {
  plot <- ggplot(data, aes(x = Time)) +
    geom_line(aes(y = Value, color = "Actual"), show.legend = FALSE) +
    geom_line(aes(y = model_1, color = "Fitted"), alpha = 0.6, show.legend = FALSE) +
    add_fitted_points_func(list_of_dfs, color_string, # make geom_point layers
                                      c("TRUE" = 1, "FALSE" = 16)) + 
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
    scale_color_manual(values = c("Actual" = "#4682B4", "Fitted" = "#FF8C00")) +
    scale_x_datetime(limits = as.POSIXct(c(start_date, end_date)), 
                     date_labels = "%m-%d",
                     #date_breaks = "3 days",
                     expand = c(0.02, 0.02)) +
    labs(title = paste(""), x = NULL, y = NULL, color = "Variable") +
    theme_bw() +
    theme(legend.position = "none",
          plot.title = element_markdown(size = 10),
          plot.title.position = "panel")
  
  return(plot)
}

```
Train train/test split
```{r, warning=FALSE}
#------------Calculate Theta - Top / bottom 5% of data ------------
top_5_percent <- quantile(train_set$.resid, probs = 0.95, na.rm = TRUE)
bottom_5_percent <- quantile(train_set$.resid, probs = 0.05, na.rm = TRUE)
top_5_percent_value <- top_5_percent[[1]] 
bottom_5_percent_value <- bottom_5_percent[[1]]
top_5_percent_value ; bottom_5_percent_value 
```
# Regression Tree
```{r}
#------------Regression tree on .resid + temporal variables------------ 
tree_model <- rpart(.resid ~ Minute + Hour + Weekday + Intensity + Steps_10, 
                    data = train_set, method = "anova", 
                    cp = 0.00000, minbucket = 10)
tree_rules <- rpart.rules(tree_model, cover = FALSE, digits = 4, nn = FALSE)
#Attached the predicted tree nodes to DF and Maps rules/leafs to Observations

leaf_nodes <- predict(tree_model, newdata = heart_merged_clean, type = "vector") 
heart_merged_with_leaf <- dplyr::bind_cols(heart_merged_clean, Leaf_Node = leaf_nodes)
```
Lift 
```{r}
train_results <- tree_alert_compute_lift(
  train_set, tree_model, "Leaf_Node", "Baseline (Train)",
  plot = TRUE, n_groups = 20,
  lift_label = "First Quantile",
  axis_title_size = 20, axis_text_x_size = 18, axis_text_y_size = 20
)
test_results <- tree_alert_compute_lift(
  test_set, tree_model, "Leaf_Node", "Baseline (Test)",
  plot = TRUE, n_groups = 20,
  lift_label = "First Quantile",
  axis_title_size = 20, axis_text_x_size = 18, axis_text_y_size = 20
)
```

## Grid Search: Joint Optimization of Tree Parameters + Predictor Variables  ##
```{r}
## ============================================================================
## GRID SEARCH: Optimize Tree Parameters ONLY
## Fixed Predictors:
## Hour + Minute + Weekday + Intensity + Steps_10
## Using SAME Lift Metric
## ============================================================================

# 1. Fixed formula (NO variable combinations)
fixed_formula <- .resid ~ Hour + Minute + Weekday + Intensity + Steps_10

# 2. Parameter grid (ONLY tuning tree hyperparameters)
param_grid <- expand.grid(
  cp        = c(0, 0.0001, 0.001, 0.005, 0.01),
  minbucket = c(5, 10, 20, 30),
  maxdepth  = c(5, 10, 15, 20, 25, 30)
)

# RUN GRID SEARCH
grid_search <- tree_alert_grid_search(
  train_set = train_set,
  formula = fixed_formula,
  param_grid = param_grid,
  n_groups = 20,
  pred_col_name = "Pred_Temp"
)

# SELECT BEST MODEL
results <- grid_search$results
best_params <- grid_search$best_params

cat("\n=============================\n")
cat("  BEST TREE PARAMETERS FOUND\n")
cat("=============================\n")
print(best_params)
```
Train Final Model using Best Parameters
```{r}
###############################################################################
## Train Final Model using Best Parameters + Best Variable Combination
###############################################################################

# Build formula dynamically from best_params$vars
final_formula <- fixed_formula

cat("  FINAL MODEL FORMULA USED\n")
print(final_formula)

# Train final rpart model
best_tree_model <- tree_alert_train_best_model(
  train_set = train_set,
  formula = final_formula,
  best_params = best_params
)

# Display tree rules
cat("\n=============================\n")
cat("        FINAL TREE RULES\n")
cat("=============================\n")
tree_rules <- rpart.rules(best_tree_model, cover = FALSE, digits = 4, nn = FALSE)
#print(tree_rules)

# Attach predicted leaf node averages to full dataset
leaf_nodes <- predict(best_tree_model, newdata = heart_merged_clean, type = "vector")
heart_merged_with_leaf <- dplyr::bind_cols(heart_merged_clean, Leaf_Node = leaf_nodes)

# Preview result
cat("\n=============================\n")
cat(" ATTACHED LEAF NODE PREDICTIONS\n")
cat("=============================\n")
print(head(heart_merged_with_leaf))
```
# LIFT CHARTS FOR BEST GRID SEARCH MODEL (Train / Test / Full)
```{r}
# ---- Compute lift for BEST MODEL ----
best_train <- tree_alert_compute_lift(
  train_set, best_tree_model, "Pred_Best", "Best Model (Train)",
  plot = TRUE, n_groups = 20,
  lift_label = "First Ventile",
  axis_title_size = 20, axis_text_x_size = 18, axis_text_y_size = 20
)
ggsave(file.path(heart_image_dir, "winkler_lift_train.png"), plot = last_plot(), width = 8, height = 6, dpi = 1000)
best_test <- tree_alert_compute_lift(
  test_set, best_tree_model, "Pred_Best", "Best Model (Test)",
  plot = TRUE, n_groups = 20,
  lift_label = "First Ventile",
  axis_title_size = 20, axis_text_x_size = 18, axis_text_y_size = 20
)
ggsave(file.path(heart_image_dir, "winkler_lift_test.png"), plot = last_plot(), width = 8, height = 6, dpi = 1000)
```
Map rules to observations and  pull top 10 rules
```{r}
start_row <- 1
end_row   <- 3

# ------------Pull top rules by leaf node value------------
top_leaf_selection <- tree_alert_top_leaf_values(
  data = heart_merged_with_leaf,
  n_rules = end_row,
  train_end_time = min(test_set$Time)
)
leaf_summary <- top_leaf_selection$leaf_summary
top_k_leaf_values <- top_leaf_selection$top_leaf_values
```

# THE PROGRAM -- appllies rules as filter and plots the results
```{r, warning=FALSE, message=FALSE}
#------------Filter the rows for each rule for plotting------------
list_of_dfs <- list() # Create an empty DataFrame to store the results
for (i in start_row:end_row) {

  filtered_data <- heart_merged_with_leaf %>%
    as.data.frame() %>%
    filter(Leaf_Node == top_k_leaf_values[i])
  
  # Find the rule row that matches the leaf node value
  rule_row_index <- which.min(abs(as.numeric(tree_rules[, 1]) - top_k_leaf_values[i]))  # <-- fix
  first_row <- tree_rules[rule_row_index, ]                                              # <-- fix
  
  non_blank_cells <- first_row[first_row != ""]
  constructed_sentence <- paste(names(non_blank_cells), non_blank_cells, collapse = " ")
  rule_text <- textGrob(constructed_sentence, gp = gpar(col = "black", fontsize = 9))
  print(constructed_sentence)
  
  #------------Plot of filtered rules on fitted time scale------------
  post_rules_scatter_plot <- ggplot(filtered_data, 
                                    aes(x = Time, y = .resid)) + 
    geom_point(aes(color = .resid > 10)) +
    geom_text(aes(label = format(Time, "%H:%M")), 
              hjust = -0.1, vjust = 0.5, size = 2) +  # Adding time labels
    labs(title = "Scatter Plot of Residuals", x = "Time", y = "Residual") +
      scale_color_manual(values = c("FALSE" = "#FF0000", "TRUE" = "#9400D3")) +
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 2)) 
  #------------Create Residual Plot------------
  #Initiate arguments
  #shape_number <- ifelse(row <= 25, row, 20)
  random_color <- rgb(runif(1), runif(1), runif(1), maxColorValue = 255)
  list_of_dfs[[i]] <- filtered_data # used to save df's for lower plot.
  # Plot of filtered rules on full series
  post_rules_line_plot <- ggplot(heart_merged_clean, 
                                 aes(x = Time, y = .resid)) +
    geom_line(color = "#00BFC4") +
    geom_point(data = filtered_data, 
               aes(x = Time, y = .resid,
                   color = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
                   shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value),
                   size = 2.5, stroke = 0.8) +
    scale_color_manual(values = c("FALSE" = color_string[i], 
                                  "TRUE" = color_string[i])) +  
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +
    labs(title = paste("Residual Plot with Tree Filtered Extreme and Non-extreme Values",
                       sep=""), x = NULL, y = NULL) +
    geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", linetype = "dashed") +
    theme_bw() +
    theme(legend.position = "none") + 
    theme(plot.title = element_markdown())
    #------------Plot of Actual Vs Predicted------------
  actual_predicted_plot <- ggplot(heart_merged_clean, aes(x = Time)) +
    geom_line(aes(y = Value, color = "Actual"), show.legend = FALSE) +
    geom_point(data = filtered_data, 
               aes(x = Time, y = Value, 
      shape = .resid >= top_5_percent_value | .resid <= bottom_5_percent_value,
      color = .resid >= top_5_percent_value | .resid <=bottom_5_percent_value),
                   size = 2.5, stroke = 1, show.legend = FALSE) +  
    scale_color_manual(values = c("TRUE" = color_string[i], 
                                  "FALSE" = color_string[i], 
                                  "Actual" = "#4682B4"),
                       guide = FALSE) + 
    scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 16)) +  
    labs(title = "<span style='color:#4682B4'><b>Actual</b>
         </span> Heart Rates + Tree Filtered Points",
         x = NULL, y = NULL, color = "Variable") +
    theme(legend.position = "none") +
    geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", linetype = "dashed") +
    theme_bw() +
    theme(plot.title = element_markdown())

   # Output plots
    grid.arrange(rule_text, post_rules_scatter_plot, ncol = 1)
    grid.arrange(post_rules_line_plot, actual_predicted_plot,
                 ncol = 1, heights = c(10,10))
}
```

```{r, warning=FALSE, message=FALSE}
#fills gaps with NA's so that the line does not appear on blanks. 
heart_merged_clean_filled <- heart_merged_clean %>%
  complete(Time = seq(min(Time), max(Time), by = "10 min"))
# -----------
# Filter points based on your criteria
highlight_points <- subset(heart_merged_clean_filled,
             (Hour %in% c("5", "6") & Weekday %in% 
                c("1", "2", "3", "4", "5")) |  # Weekday early AM
             Weekday %in% c("7") |  # Weekends
             WeekofMonth %in% c("1", "4", "5")) # 1st/4th/5th weeks of the month

# Plot of filtered rules on full series
post_rules_line_plot <- ggplot(heart_merged_clean_filled, aes(x = Time, y = .resid)) +
  geom_line(color = "#00BFC4") +
  geom_point(data = highlight_points, 
             aes(x = Time, y = .resid, color = .resid > 10), size = 0.5) +  
  labs(title = "Residual Plot: <span style='color:#9900D0'><b>Extreme</b></span> vs <span 
       style='color:#F8666D'><b>Non-extreme</b></span> values",
       x = NULL, y = NULL, color = "Variable") +
  scale_color_manual(values = c("FALSE" = "#F8766D", "TRUE" = "#C77CFF")) +
  scale_x_datetime(date_breaks = "4 day", date_labels = "%m-%d", 
                   expand = c(0.02, 0.02)) +
  theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown())

#------------All Filters Actual Vs Predicted------------
# Create the plot
actual_predicted_plot <- ggplot(heart_merged_clean_filled, aes(x = Time)) +
  geom_line(aes(y = Value, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  geom_line(aes(y = .fitted, color = "Fitted"), 
            alpha = 0.6, show.legend = FALSE) +
  add_fitted_points_func(list_of_dfs, color_string, 
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 1)) +
  scale_color_manual(values = c("Actual" = "#4682B4", "Fitted" = "#FF8C00")) +
  labs(title = "<span style='color:#4682B4'><b>Actual</b></span> & 
       <span style='color:#FF8C00'><b>Fitted</b></span> Heart Rates", 
       x = NULL, y = NULL, color = "Variable") +
  geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", 
             linetype = "dotted", size = 1) +
  theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10))

actual_predicted_plot

# Create the plot
residual_plot <- ggplot(heart_merged_clean_filled, aes(x = Time)) +
  geom_line(aes(y = .resid, color = "Actual"),  na.rm = FALSE, show.legend = FALSE) +
  add_resid_points_func(list_of_dfs, color_string, # Generate geom_point layers
                                      c("TRUE" = 1, "FALSE" = 16)) + 
  scale_shape_manual(values = c("TRUE" = 1, "FALSE" = 1)) +
  scale_color_manual(values = c("Actual" = "#00BFC4")) +
  # Change Title to from errors to Winkler Score / ect when option 2 is selected
  labs(title = "Errors", x = NULL, y = NULL, color = "Variable") +   
  geom_vline(xintercept = min(test_set$Time, na.rm = TRUE), color = "red", 
             linetype = "dotted", size = 1) + theme_bw() +
  theme(legend.position = "none") + 
  theme(plot.title = element_markdown(size = 10)) 
residual_plot

two_plots <- grid.arrange(actual_predicted_plot, residual_plot, ncol = 1)
# Export the plot
#ggsave("HeartRate.png", 
#       plot = two_plots, width = 6, height = 4, dpi = 800) 
ggsave(file.path(heart_image_dir, "hr_viz.png"), 
       plot = two_plots, width = 6, height = 4.5, dpi = 800)
```
Misc
```{r}
percent_above_10 <- mean(highlight_points$.resid > 10) * 100
average_resid <- mean(highlight_points$.resid)
cat("Percentage of values above 10:", percent_above_10, "%\n")
cat("Average value of .resid column:", average_resid, "\n")
```
