---
title: "TreeAlert Ordered Lift Validation: Electricity"
author: "Bobea"
date: "2024-02-03"
output: html_document
---

## Purpose

This standalone validation evaluates the stability of the electricity-demand TreeAlert lift result. The chronological training sample is divided into 20 contiguous blocks. One block is removed at a time, the tree hyperparameters are re-optimized on the remaining training data, and lift is calculated for the retained training sample and the unchanged external test sample.

## Libraries and Configuration

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(rpart)
library(scales)
library(tidyr)

project_root <- if (dir.exists(file.path("..", "TA_functions"))) ".." else "."
source(file.path(project_root, "TA_functions", "TA_functions.R"))

validation_results_dir <- file.path(project_root, "Validation", "results")
validation_images_dir <- file.path(project_root, "Validation", "images")
dir.create(validation_results_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(validation_images_dir, showWarnings = FALSE, recursive = TRUE)

Actual_column <- "DE_LU_load_actual_entsoe_transparency"
Forecast_column <- "DE_LU_load_forecast_entsoe_transparency"

data_start <- "2019-02-01"
data_end <- "2019-5-30"
train_start <- "2019-02-01"
train_end <- "2019-5-01"
test_start <- train_end
test_end <- data_end

n_folds <- 20
n_groups <- 20
lift_metric <- "mean_abs"
validation_formula <- .resid ~ Hour + Minute + Weekday

param_grid <- expand.grid(
  cp        = c(0, 0.0001, 0.001, 0.005, 0.01),
  minbucket = seq(4, 8, by = 1),
  maxdepth  = seq(4, 8, by = 1),
  stringsAsFactors = FALSE
)
```

## Data Preparation

```{r data-preparation, warning=FALSE, message=FALSE}
temporal_encoding <- function(data) {
  data <- data %>%
    mutate(
      Time = parse_date_time(Time, orders = c("ymd HMS", "ymd HM", "ymd H", "ymd")),
      Hour = factor(hour(Time)),
      Minute = factor(minute(Time)),
      Weekday = factor(lubridate::wday(Time, label = TRUE)),
      DayofMonth = factor(mday(Time)),
      WeekofMonth = factor(ceiling(day(Time) / 7)),
      Month = factor(lubridate::month(Time, label = TRUE, abbr = TRUE))
    ) %>%
    arrange(Time)

  for (week in 1:4) {
    lag_label <- paste0("Lag_", week, "_Week")
    lag_time_col <- paste0("LagTime_", week, "_Week")

    data <- data %>%
      mutate(!!lag_time_col := Time - weeks(week)) %>%
      left_join(
        data %>%
          select(Time, .resid) %>%
          rename(!!lag_label := .resid),
        by = setNames("Time", lag_time_col)
      ) %>%
      select(-all_of(lag_time_col))
  }

  data %>%
    arrange(Time) %>%
    fill(starts_with("Lag_"), .direction = "down")
}

process_data <- function(data, time_col, actual_col, forecast_col) {
  data %>%
    select({{ time_col }}, {{ actual_col }}, {{ forecast_col }}) %>%
    rename(
      Time = {{ time_col }},
      actual = {{ actual_col }},
      forecast = {{ forecast_col }}
    ) %>%
    mutate(
      Time = parse_date_time(
        Time,
        orders = c("ymd HMS", "ymd HM", "mdy HMS", "mdy HM", "dmy HMS", "dmy HM")
      ),
      .resid = actual - forecast
    ) %>%
    na.omit()
}

raw_data <- read.csv(file.path(
  project_root,
  "Electricity",
  "Elec_data",
  "time_series_15min_singleindex.csv"
)) %>%
  slice(-(1:5))

analysis_data <- process_data(
  raw_data,
  time_col = utc_timestamp,
  actual_col = Actual_column,
  forecast_col = Forecast_column
)

analysis_data <- analysis_data[
  analysis_data$Time >= as.POSIXct(data_start, tz = "UTC") &
    analysis_data$Time < as.POSIXct(data_end, tz = "UTC"),
]

data_filtered <- analysis_data %>%
  mutate(hour = format(Time, "%Y-%m-%d %H")) %>%
  group_by(hour) %>%
  filter(actual == max(actual)) %>%
  ungroup() %>%
  select(-hour)

analysis_data <- temporal_encoding(data_filtered) %>%
  mutate(
    Value = actual,
    .fitted = forecast,
    .resid = .resid
  ) %>%
  select(-actual, -forecast) %>%
  as_tibble()

train_set <- analysis_data[
  analysis_data$Time > as.POSIXct(train_start, tz = "UTC") &
    analysis_data$Time < as.POSIXct(train_end, tz = "UTC"),
]
test_set <- analysis_data[
  analysis_data$Time >= as.POSIXct(test_start, tz = "UTC") &
    analysis_data$Time < as.POSIXct(test_end, tz = "UTC"),
]
```

## Ordered Leave-One-Block-Out Validation

```{r ordered-validation, warning=FALSE, message=FALSE}
validation_results <- tree_alert_ordered_leave_block_validation(
  data = train_set,
  test_set = test_set,
  formula = validation_formula,
  param_grid = param_grid,
  n_folds = n_folds,
  n_groups = n_groups,
  pred_col_name = "Pred_Electricity_Validation",
  cp_digits = 6,
  lift_metric = lift_metric
)

validation_plot <- tree_alert_ordered_validation_plot(
  validation_results,
  image_path = file.path(
    validation_images_dir, "electricity_ordered_lift_validation.png"
  ),
  legend_position = c(0.98, 0.98),
  legend_justification = c("right", "top")
)

write.csv(
  validation_results$fold_results,
  file.path(validation_results_dir, "electricity_ordered_lift_folds.csv"),
  row.names = FALSE
)
write.csv(
  validation_results$interval_summary,
  file.path(validation_results_dir, "electricity_ordered_lift_intervals.csv"),
  row.names = FALSE
)

validation_results$interval_summary
```

```{r plot, echo=FALSE, fig.width=8.5, fig.height=5.5, dpi=300}
validation_plot
```

The shaded bands are empirical 95% intervals from the 20 ordered folds. The intervals are interpreted relative to the lift-one benchmark rather than as a formal hypothesis test. An interval whose lower bound exceeds one indicates that the observed fold-level lift remains above the no-concentration benchmark across the empirical interval.
