---
title: "TreeAlert Ordered Lift Validation: Heart Rate"
author: "Bobea"
date: "2024-02-03"
output: html_document
---

## Purpose

This legacy ordered leave-block validation is retained for reference only and is no longer executed. The TreeAlert boxplots now use randomized 20-fold cross-validation, with each fold evaluated on its held-out training observations and on the fixed external test set.

## Libraries and Configuration

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(rpart)
library(scales)

project_root <- if (dir.exists(file.path("..", "TA_functions"))) ".." else "."
source(file.path(project_root, "TA_functions", "TA_functions.R"))

validation_results_dir <- file.path(project_root, "Validation", "results")
validation_images_dir <- file.path(project_root, "Validation", "images")
dir.create(validation_results_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(validation_images_dir, showWarnings = FALSE, recursive = TRUE)

n_folds <- 20
n_groups <- 20
lift_metric <- "rmse"
validation_formula <- .resid ~ Hour + Minute + Weekday + Intensity + Steps_10

param_grid <- expand.grid(
  cp        = c(0, 0.0001, 0.001, 0.005, 0.01),
  minbucket = seq(4, 8, by = 1),
  maxdepth  = seq(1, 6, by = 1)
)
```

## Data Preparation

```{r data-preparation, warning=FALSE, message=FALSE}
heart_data <- readRDS(file.path(
  project_root, "Heart_Rate", "HR_data", "heart_combined.rds"
))

split_index <- floor(0.85 * nrow(heart_data))
train_set <- heart_data[seq_len(split_index), , drop = FALSE]
test_set <- heart_data[(split_index + 1):nrow(heart_data), , drop = FALSE]
```

## Deprecated Ordered Leave-One-Block-Out Validation (not executed)

```{r ordered-validation, eval=FALSE, warning=FALSE, message=FALSE}
validation_results <- tree_alert_ordered_leave_block_validation(
  data = train_set,
  test_set = test_set,
  formula = validation_formula,
  param_grid = param_grid,
  n_folds = n_folds,
  n_groups = n_groups,
  pred_col_name = "Pred_Heart_Rate_Validation",
  lift_metric = lift_metric
)

validation_plot <- tree_alert_ordered_validation_plot(
  validation_results,
  image_path = file.path(
    validation_images_dir, "heart_rate_ordered_lift_validation.png"
  ),
  legend_position = c(0.02, 0.12),
  legend_justification = c("left", "bottom")
)

write.csv(
  validation_results$fold_results,
  file.path(validation_results_dir, "heart_rate_ordered_lift_folds.csv"),
  row.names = FALSE
)
write.csv(
  validation_results$interval_summary,
  file.path(validation_results_dir, "heart_rate_ordered_lift_intervals.csv"),
  row.names = FALSE
)

validation_results$interval_summary
```

```{r plot, eval=FALSE, echo=FALSE, fig.width=8.5, fig.height=5.5, dpi=300}
validation_plot
```

<!-- Legacy ordered-validation interpretation disabled. Active lift distributions are generated by randomized 20-fold cross-validation in the main analyses and Appendix D. -->
